---
title: "Comparison between sequences at 100% identity"
subtitle: "References for taxonomical assignation: PR2 vs MareChiara"
author: "Lucia Campese"
date: "September 2020"
output:
  html_document:
    code_folding: hide
---
---
  
  ```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<style>
.column-left{
  float: left;
  width: 50%;
  text-align: left;
}
.column-right{
  float: right;
  width: 50%;
  text-align: right;
}
</style>
  
<br>

### Aim
The Aim is to compare distribution, biogeography and taxonomic composition of the metabarcoding V4 dataset taxonomically assigned to the genus *Pseudonitzschia* using two different reference databases:

a) A revised version  of the **PR2 database** (Guillouet al., 2013) for which the Chlorophyta sequences  had  been  checked  against  the  latest taxonomy (Traginet al., 2016) [info to be confirmed] => **TARA-PR2** ; 

b) A selection of 27 *Pseudonitzschia* species sequences (from Valeria R.) from **MareChiara** V4 dataset => **TARA-MC**

<br>

### Dataset
##### Data: TARA Oceans and Tara Oceans Polar circle metabarcording (V4)
##### Taxon: *Pseudonitzschia*
##### Geographic scale: Global


```{bash, eval = FALSE}
#1. Extract Pseudonitzschia sequences from the total V4 dataset

pwd: ~/Data/metaB/V4/TARA_exploration/TARA_identity100

srun grep -Ff Pn_sequenceId.txt ~/Data/metaB/V4/18S_V4_1191_samples.results > 18S_V4_1191_samples_Pn.results &

###30946 sequences
```


### Workflow

<br>

#### **Building datasets**

<br>

##### a) **TARA-PR2**

1. Check the distribution of the identity percentage of amplicons to the reference dataset (PR2)

```{r warning=FALSE, message=FALSE, cache=TRUE, fig.align='center'}

library(data.table)
library(tidyverse)
library(gridExtra)


load("C:/Users/Userszn/Documents/PhD/script/Pseudonitzschia/metaB/V4/Pn_tax.RData")

Percid <- 
  Pn_tax %>% 
  select(perc.identity) %>% 
  summary()


mainplot <-Pn_tax %>% ggplot + 
  aes(x = perc.identity, 
      fill = tot.numb.reads) + 
  geom_histogram(bins = 100) + 
  theme_bw() + 
  # xlim(90,100) +
  annotation_custom(tableGrob(Percid), xmin=60, xmax=70, ymin=5000, ymax=12000) +
  ggtitle("Percentage of identity to the PR2 reference", 
          subtitle= paste0("Pseudonitzschia, TARA V4; ", 
                           n_distinct(Pn_tax$seqId), 
                           " distinct amplicons")) +
  annotate("rect", xmin=95, xmax=100, ymin=0, ymax=12000, fill="lightblue", alpha = 0.5)


zoomplot <- Pn_tax %>% ggplot + 
  aes(x = perc.identity, 
      fill = tot.numb.reads) + 
  geom_histogram() + 
  theme_bw() + 
  xlim(95,100) +
  ggtitle("zooming") +
  #theme(panel.background = element_rect(fill = 'lightblue', colour = 'lightblue')) +
  theme(plot.background = element_rect(fill = 'lightblue'))



mainplot + annotation_custom(ggplotGrob(zoomplot), xmin = 75, xmax = 90, 
                             ymin = 5000, ymax = 11000)
```


The identity percentage ranges from 59.8 to 100, but almost all sequences show a value around 99% identity to the reference database.

<br>

2.	Extract from the dataset only the sequences with percentage of identity to PR2 database of 100% 

```{R, eval = FALSE}

#take only sequences with at least 3 reads (seen three times in the whole dataset) [*]
srun awk '$2>2' 18S_V4_1191_samples_Pn.results > 18S_V4_1191_samples_Pn_3readsmin.results &

#wc -l: 25453 18S_V4_1191_samples_Pn_3readsmin.results
#wc -l: 30946 18S_V4_1191_samples_Pn.results


#[*]as adviced by Nicholas Henry.
  
  
#Take only 100% pid sequences (pid is 3rd column)
awk '$3>99.99' 18S_V4_1191_samples_Pn_3readsmin.results > 18S_V4_1191_samples_Pn_pid100_3readsmin.results

#they are 17

#Extract only seq ids (column 1)
awk '{print $1}' 18S_V4_1191_samples_Pn_pid100_3readsmin.results > 18S_V4_1191_samples_Pn_pid100_3readsmin_seqid.txt

#use the retrieved 17 sequence Ids to subset the distr file (i.e. the one 
#containing seqId|sampleId|readcounts

srun grep -Ff 18S_V4_1191_samples_Pn_pid100_3readsmin_seqid.txt ~/Data/metaB/V4/18S_V4_1191_samples.distr > 18S_V4_1191_samples_Pn_pid100_3readsmin.distr &


#the final output is made by 17 different sequences

```

There are **17** different **sequences** assigned to Pseudonitzschia species with a **100% identity** to PR2 database and occurring at least **3 times** in the dataset. 

<br>

3. Add the environmental information for the selected 17 sequences

Note: All sequences have length >380 bp [[[it's sequence length and not wuery cover?]]]:


```{r warning=FALSE, message=FALSE, cache=TRUE}
library(data.table)
library(tidyverse)

Pn_identity100.distr <- fread("C:/Users/Userszn/Documents/PhD/script/Pseudonitzschia/metaB/V4/18S_V4_1191_samples_Pn_pid100_3readsmin.distr") #14 sequences

Pn_identity100.tax <- fread("C:/Users/Userszn/Documents/PhD/script/Pseudonitzschia/metaB/V4/18S_V4_1191_samples_Pn_pid100_3readsmin.results", header = F, sep = "\t") #17 sequences

###ADD LENGTH INFO
length.info <-fread("C:/Users/Userszn/Documents/PhD/script/Pseudonitzschia/metaB/V4/18S_V4_1191_samples_Pn_pid100_3readsmin_seqid_length_seq_clean")



###ADD LENGTH INFO
length.info <-fread("C:/Users/Userszn/Documents/PhD/script/Pseudonitzschia/metaB/V4/18S_V4_1191_samples_Pn_pid100_3readsmin_seqid_length_seq_clean")

Pn_identity100.tax.length <- length.info %>% 
  rename(length = V2) %>% 
  left_join(Pn_identity100.tax)

Pn_identity100.tax.length <- length.info %>% 
  rename(length = V2) %>% 
  left_join(Pn_identity100.tax)


Pn_identity100.distr.tax.length <- inner_join(Pn_identity100.distr, Pn_identity100.tax.length, by = "V1") %>% 
  select(-V2.y)

colnames(Pn_identity100.distr.tax.length) <- c("seqId", "sample", "readcounts", "length", "perc.identity", "lineage", "idpr2")


env <- fread("C:/Users/Userszn/Documents/PhD/script/Pseudonitzschia/metaB/V4/18S_V4_1191_samples_context.tsv")

Pn_identity100.distr.tax.env <- inner_join(Pn_identity100.distr.tax.length, env)


Pn_identity100.distr.tax.env1 <- splitstackshape::cSplit(Pn_identity100.distr.tax.env, "lineage", "|") %>% 
  select(seqId:minimum.generation.time.h,lineage_13) %>% 
  mutate(species = case_when(lineage_13 == "*" ~ "Pseudo-nitzschia+unknown",
                             TRUE ~ as.character(lineage_13))) %>% 
  select(-lineage_13) %>% 
  splitstackshape::cSplit("species", "+") %>%
  rename(species = species_2) %>% 
  group_by(seqId) %>% 
  ungroup() %>% 
  select(seqId, length, sample, readcounts, perc.identity:idpr2, species, Station.label:Depth.chloro.max)


final_tara_pr2<- Pn_identity100.distr.tax.env1 %>% 
  select(seqId, length, species, sample, readcounts, Station.label:Depth.chloro.max) %>%
  mutate(dataset = c("TARA-PR2"))


#final_taraPR2 is the final file with seqIds, taxonomical annotation, distribution and environmental info + LENGTH
```

<div class="column-left">
```{r warning=FALSE, message=FALSE, cache=TRUE}
distr.lengths <- final_tara_pr2 %>% 
  select(length) %>% 
  rename('sequence length' = length) %>% 
  summary()


library(kableExtra)

distr.lengths %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```
</div><div class="column-right">
```{r warning=FALSE, message=FALSE, cache=TRUE}
final_tara_pr2 %>% 
  select(seqId, length) %>% 
  unique() %>% 
  ggplot() + 
  aes(x = length) + 
  geom_histogram(bins = 100, stat = "count") + 
  theme_bw() 
```
</div>

<div class="column-left">
```{r warning=FALSE, message=FALSE, cache=TRUE}
distr.lengths <- final_tara_pr2 %>% 
  select(length) %>% 
  rename('sequence length' = length) %>% 
  summary()


library(kableExtra)

distr.lengths %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```
</div><div class="column-right">
```{r warning=FALSE, message=FALSE, cache=TRUE}
final_tara_pr2 %>% 
  select(seqId, length) %>% 
  unique() %>% 
  ggplot() + 
  aes(x = length) + 
  geom_histogram(bins = 100, stat = "count") + 
  theme_bw() 
```
</div>

<br>


```{r eval=FALSE, message=FALSE, cache=TRUE}
print("ue")
```

##### b) **TARA-MC**

<br>

1. Extraction of sequences annotated at 100% with the reference:

[length to be checked]

```{bash, eval = FALSE}

pwd: /home/lcampese/Data/metaB/V4/Blast_MC_TO/identity100

#remove the "size" part after the id of each sequence
cut -f1 -d";" Pn_tara_candi_final_pid100_list.txt > Pn_tara_candi_final_pid100_list_clean.txt


#508 sequenze 

#subset the results file using my seq Ids
srun grep -Ff Pn_tara_candi_final_pid100_list_clean.txt ~/Data/metaB/V4/TARA_exploration/18S_V4_1191_samples_Pn_3readsmin.results > 18S_V4_1191_samples_Pncandi_3readsmin_pid100.results &

#324 sequenze

#extract the ids for these 324 seq
awk '{print $1}' 18S_V4_1191_samples_Pncandi_3readsmin_pid100.results > Pn_tara_candi_final_pid100_3reads_list_clean.txt

#subset the distr file using these 324 seqIds

srun grep -Ff Pn_tara_candi_final_pid100_3reads_list_clean.txt ~/Data/metaB/V4/18S_V4_1191_samples.distr > 18S_V4_1191_samples_Pncandi_3readsmin_pid100.distr &

```

There are 324 different sequences assigned to Pseudonitzschia species with a **100% identity** to MareChiara database and occurring at least **3 times** in the dataset. 

<br> 

2. Add the environmental information for the selected 324 sequences

```{r warning=FALSE, message=FALSE, cache=TRUE}

Pn_candi_identity100.distr <- fread("C:/Users/Userszn/Documents/PhD/script/Pseudonitzschia/metaB/V4/cfr100pid/18S_V4_1191_samples_Pncandi_3readsmin_pid100.distr") 

colnames(Pn_candi_identity100.distr) <- c("seqId", "sample", "readcounts")

paste0(n_distinct(Pn_candi_identity100.distr$seqId), " sequence Ids")

Pn_candi_identity100.tax <- fread("C:/Users/Userszn/Documents/PhD/script/Pseudonitzschia/metaB/V4/cfr100pid/Pn_tara_candi_final_pid100.txt") %>%
  separate(V1, into = c("seqId", "size"), sep = ";") %>% 
  rename(tax = V2, pid = V3) %>% 
  select(seqId, tax, pid)


#subset tax (still with 508 seq) using the distr file (with the corret 324 seq) {the aim is to add the taxonomy info}

Pn_candi_identity100.distr.tax <- inner_join(Pn_candi_identity100.tax, Pn_candi_identity100.distr, by = "seqId")

Pn_candi_identity100.distr.tax <- inner_join(Pn_candi_identity100.tax, Pn_candi_identity100.distr, by = "seqId")

##add env

env <- fread("C:/Users/Userszn/Documents/PhD/script/Pseudonitzschia/metaB/V4/18S_V4_1191_samples_context.tsv")

Pn_candi_identity100.distr.tax.env <- inner_join(Pn_candi_identity100.distr.tax, env) %>% 
  select(seqId:readcounts,Pangaea_id:minimum.generation.time.h)


final_tara_mc <- Pn_candi_identity100.distr.tax.env %>% 
  mutate(species = case_when(grepl("allochrona", tax) ~ "allochrona",
                             grepl("americana", tax) ~ "ID2_amer|cacia|circump|granii",
                             grepl("arenysensis", tax) ~ "ID1_aren|dolo|arc|subcur",
                             grepl("ID3_P._australis", tax) ~ "ID3_australis|multiseries",
                             grepl("P._australis_SPC21", tax) ~ "australis_single",
                             grepl("batesiana", tax) ~ "batesiana",
                             grepl("brasiliana", tax) ~ "ID4_brasiliana|lineola",
                             grepl("calliantha", tax) ~ "calliantha",
                             grepl("cuspidata", tax) ~ "cuspidata",
                             grepl("decipiens", tax) ~ "decipiens",
                             grepl("delicatissima", tax) ~ "delicatissima",
                             grepl("fraudulenta", tax) ~ "fraudulenta",
                             grepl("fukuyoi", tax) ~ "fukuyoi",
                             grepl("galaxiae", tax) ~ "galaxiae",
                             grepl("heimii", tax) ~ "heimii",
                             grepl("kodamae", tax) ~ "ID5_kodamae|lundholmiae",
                             grepl("mannii", tax) ~ "mannii",
                             grepl("micropora", tax) ~ "micropora",
                             grepl("multistriata", tax) ~ "multistriata",
                             grepl("pseudodelicatissima", tax) ~ "pseudodelicatissima",
                             grepl("pungens", tax) ~ "pungens",
                             grepl("sabit", tax) ~ "sabit",
                             grepl("turgidula", tax) ~ "turgidula",
                            TRUE ~ tax)) %>% 
  mutate(length = "tobedone") %>% 
  select(seqId, length, species, sample, readcounts, Station.label:Depth.chloro.max) %>%
  mutate(dataset = c("TARA-MC"))
```

<br>

<br>


#### **Comparison** 

<br>

1. How big is the overlap between the two obtained datasets?

<div class="column-left">
```{r warning=FALSE, message=FALSE, cache=TRUE, fig.height=4, fig.width=4, fig.align="center"}

#Sequences in common: Venn Diagram

tara_pr2_venn <- final_tara_pr2 %>% 
  select(seqId) %>% 
  mutate(`TARA-PR2` = c("1")) %>% 
  unique()

tara_mc_venn <- final_tara_mc %>% 
  select(seqId) %>% 
  mutate(`TARA-MC` = c("1")) %>% 
  unique()


print("Sequence Ids in common")

full_join(tara_pr2_venn, tara_mc_venn) %>% 
  replace_na(list(`TARA-PR2` = 0, `TARA-MC` = 0)) %>% 
  select(-seqId) %>%
  venn::venn(ellipse = F, borders = F,  cexil = 0.8, cexsn = 0.9,
             ilcs = 0.8, sncs = 0.8, box = F, col = "black",
             zcolor = "skyblue, lightgreen")
```
</div><div class="column-right">
```{r warning=FALSE, message=FALSE, cache=TRUE, fig.height=4, fig.width=4, fig.align="center"}

###species in common
tara_pr2_venn_species <- final_tara_pr2 %>% 
  select(species) %>% 
  mutate(`TARA-PR2` = c("1")) %>% 
  unique()

tara_mc_venn_species <- final_tara_mc %>% 
  select(species) %>% 
  mutate(`TARA-MC` = c("1")) %>% 
  unique()


print("Species in common")

full_join(tara_pr2_venn_species, tara_mc_venn_species) %>% 
  replace_na(list(`TARA-PR2` = 0, `TARA-MC` = 0)) %>% 
  select(-species) %>% 
  venn::venn(ellipse = F, borders = F,  cexil = 0.8, cexsn = 0.9,
             ilcs = 0.8, sncs = 0.8, box = F, col = "black",
             zcolor = "skyblue, lightgreen")
```
</div>

<br>

The two datasets are quite different in the number of amplicons. **TARA-PR2** has **17 sequences**, while **TARA-MC** has **324**. There are **8 shared sequences** between the two. In terms of species, they share 4 species, with TARA-PR2 containing 3 species never detected in TARA-MC, and the latter with 17 species never annotated in TARA-PR2. 


```{r warning=FALSE, message=FALSE, cache=TRUE, fig.align="center"}

plot1 <- rbind(final_tara_pr2, final_tara_mc) %>%  
  group_by(dataset) %>% 
  summarise(tot.numb.reads = sum(readcounts)) %>% 
  ggplot() +
  aes(x=dataset, y=tot.numb.reads, fill = dataset) +
  geom_bar(stat = "identity", alpha = 0.5) +
  theme_bw() +
  scale_fill_manual(values = c("TARA-MC" = "lightgreen",
                               "TARA-PR2" = "skyblue")) +
  xlab("") +
  ylab("total number of read counts")

plot2 <-  rbind(final_tara_pr2, final_tara_mc) %>% 
  mutate(Station.label = substr(Station.label, 6, 8)) %>% 
  group_by(seqId, dataset) %>% 
  summarise(tot.numb.reads = sum(readcounts)) %>% 
  ggplot() +
  geom_histogram(aes(tot.numb.reads, fill=dataset), alpha=0.5) +
  theme(legend.position = "none") +
  theme_bw() +
  scale_x_log10() +
  xlab("read counts per sequence Id") +
  scale_fill_manual(values = c("TARA-MC" = "lightgreen",
                               "TARA-PR2" = "skyblue")) 

library(patchwork)

((plot1 | plot2) & theme(legend.position = "bottom")) + plot_layout(guides = "collect")
```


The total number of reads (plot on the left) is higher than 400000 in both datasets, with TARA-MC containing around 50000 of reads more than TARA-MC.

By looking at the distribution of the total number of reads per each amplicon (plot on the right), we can notice that TARA-MC has a high number of sequences with few read counts, and few very abundant sequences. This is not evident for TARA-PR2, that displays a more eterogeneous pattern. 

<br>

2. How is the distribution of **richness** and **abundance** per **sampling station**?


```{r warning=FALSE, message=FALSE, cache=TRUE, fig.align="center", fig.width=12, fig.height=10}

rbind(final_tara_pr2, final_tara_mc) %>%
  mutate(Station.label = substr(Station.label, 6, 8)) %>% 
  group_by(Station.label, dataset) %>% 
  summarise(tot.reads.station = sum(readcounts),
         richness.station.spp =n_distinct(species)) %>% 
  ungroup() %>% 
  pivot_longer(tot.reads.station:richness.station.spp) %>% 
  ggplot() +
  aes(x=Station.label, y=value, col=dataset, group=dataset) +
  geom_line(size=1.5) +
  theme_bw() +
  scale_color_manual(values = c("TARA-MC" = "lightgreen",
                               "TARA-PR2" = "skyblue")) +
  theme(axis.text.x = element_text(angle=50, size=6, vjust = 0.5))  +
  facet_wrap(~ name, scales = "free", nrow = 2)
```

```{r eval=FALSE}

rbind(final_tara_pr2, final_tara_mc) %>%
  mutate(Station.label = substr(Station.label, 6, 8)) %>% 
  group_by(Station.label, dataset) %>% 
  summarise(tot.reads.station = sum(readcounts)) %>% 
  ungroup() %>%
  ggplot() +
  geom_histogram(aes(tot.reads.station, fill = dataset), alpha=0.5) +
  theme_minimal() +
  theme(legend.position = "none") +
  scale_x_log10() +
  xlab("abundance per station") +
   scale_fill_manual(values = c("TARA-MC" = "lightgreen",
                               "TARA-PR2" = "skyblue")) +
  ggtitle("Pseudo-nitzschia abundance expressed as number of reads per station") +
  facet_wrap(~dataset, scales="free_y")
```



The **richness** - i.e. number of species detected - (upper plot) follows the same trend between the two datasets along the sampling stations, with few exceptions (i.e. stations 31, 47, 54). 
This trend is even more evident when looking at the total abundance per station - expressed as total number of read counts per stations- (bottom plot), where the two lines overlap for almost all samples. One exception is the case of station 155, where TARA-MC exceeds TARA-PR2 by 4000 units ca.

<br>

- Let's look at the geographical distribution of these values

```{r warning=FALSE, message=FALSE, cache=TRUE, fig.height=10, fig.width=12}

library(maps)

world_map <- map_data("world")
p <- ggplot() + coord_fixed()

base_world3 <- p + geom_polygon(data=world_map,aes(x=long, y=lat,group=group),colour="gray80",size=0.2, fill="gray80")+
  theme_bw() + coord_fixed()


#map
combined_datasets_map <- rbind(final_tara_pr2, final_tara_mc) %>%
  group_by(Station.label, dataset) %>% 
  mutate(Lat= mean(Latitude),
         Long=mean(Longitude),
         abundance.per.station = sum(readcounts),
         richness.per.station.amplicons = n_distinct(seqId),
         richness.per.station.species = n_distinct(species)) %>%
  ungroup() %>% 
  select(seqId, species, Station.label, Lat, Long, dataset, abundance.per.station,
         richness.per.station.amplicons, richness.per.station.species) %>% unique()
  
base_world3+
  geom_point(data=combined_datasets_map,
             aes(x = Long, y = Lat, 
                 size = abundance.per.station,
                 fill = richness.per.station.species),
                 pch = 21, col = "black") +
  scale_fill_viridis_c() +
  theme_minimal() +
  theme(legend.text=element_text(size=10)) +
  facet_wrap(~dataset, nrow = 2) +
  ggtitle("Pseudo-nitzschia abundance and richness per station (V4)",
        "Identity 100%") 
```


<br>

<br>

3. Is there a **correlation** between **richness** and **abundance**?


```{r warning=FALSE, message=FALSE, cache=TRUE, fig.align="center", fig.width=12, fig.height=10}
corpr2 <- rbind(final_tara_pr2, final_tara_mc) %>%
  mutate(Station.label = substr(Station.label, 6, 8)) %>% 
  group_by(Station.label, dataset) %>% 
  summarise(tot.reads.station = sum(readcounts),
            richness.station.spp =n_distinct(species)) %>% 
  ungroup() %>% 
  filter(dataset == "TARA-PR2")

corpr2.res<-psych::corr.test(corpr2$richness.station.spp, corpr2$tot.reads.station, method = "spearman")
#print(corpr2.res, short = F)

print("correlation between richness and abundance in TARA-PR2")
cor.test(corpr2$richness.station.spp, corpr2$tot.reads.station, method = "spearman")

cormc <- rbind(final_tara_pr2, final_tara_mc) %>%
  mutate(Station.label = substr(Station.label, 6, 8)) %>% 
  group_by(Station.label, dataset) %>% 
  summarise(tot.reads.station = sum(readcounts),
            richness.station.spp =n_distinct(species)) %>% 
  ungroup() %>% 
  filter(dataset == "TARA-MC")


cormc.res<-psych::corr.test(cormc$richness.station.spp, cormc$tot.reads.station, method = "spearman")
#print(cormc.res, short = F)

print("correlation between richness and abundance in TARA-MC")
cor.test(cormc$richness.station.spp, cormc$tot.reads.station, method = "spearman")

```

There is a significant correlation between richness and abundance in both dataset, weaker in TARA-PR2 (rho = 0.3) and stronger in TARA-MC (rho = 0.5).

<br>

<br>

4. Is there **spatial autocorrelation**?

<div class="column-left">
```{r warning=FALSE, message=FALSE, cache=TRUE, fig.align="center"}

print("TARA-PR2")
final_tara_pr2 %>% 
  mutate(Station.label = substr(Station.label, 6, 8)) %>% 
  group_by(Station.label, dataset) %>% 
  summarise(sum = sum(readcounts)) %>% 
  ungroup() %>%
  arrange(Station.label) %>% 
  pull(sum) %>% 
  acf
```
</div><div class="column-right">
```{r warning=FALSE, message=FALSE, cache=TRUE, fig.align="center"}
print("TARA-MC")
final_tara_mc %>% 
  mutate(Station.label = substr(Station.label, 6, 8)) %>% 
  group_by(Station.label, dataset) %>% 
  summarise(sum = sum(readcounts)) %>% 
  ungroup() %>%
  arrange(Station.label) %>% 
  pull(sum) %>% 
  acf
```
</div>


The plots show thata **there is spatial autocorrelation in TARA-PR2**. 
<<<<<<< HEAD

In particular, in TARA-PR2 graph (left) there is one spike at 1 that clearly falls outside the bonds, meaning that the total abundance in each station consistently influences the abundance in the following station.

Moreover, in both dataset there is a peak at 14 that slightly exceeds the outside bonds and one at 15 that almost does.

=======

In particular, in TARA-PR2 graph (left) there is one spike at 1 that clearly falls outside the bonds, meaning that the total abundance in each station consistently influences the abundance in the following station.

Moreover, in both dataset there is a peak at 14 that slightly exceeds the outside bonds and one at 15 that almost does.

Besides that, the abundance per station values are independent of each other if only trying to explain them with themselves (the autocorrelation property).

How to deal with that?
Spatial autocorrelation have implications when applying statistical methods (to be discussed) 

<br>

<br>


5. What about **taxonomy**?

5.1. Is the taxonomical annotation of the 8 shared sequence the same?

```{r warning=FALSE, message=FALSE, cache=TRUE}

#common sequences: is the taxonomy the same?

cfr_taxonomy_shared <- final_tara_pr2 %>%
  select(seqId, species) %>%
  rename(PR2 = species) %>% 
  inner_join(final_tara_mc[,c(1:6)]) %>%
  rename(MC = species) %>%
  select(seqId, PR2, MC) %>% 
  unique() #%>% 
 # mutate(MC = case_when(grepl("cuspidata", MC_0) ~ "cuspidata",
  #                          grepl("americana", MC_0) ~ "americana",
  #                          grepl("arenysensis", MC_0) ~ "arenysensis",
   #                         grepl("fraudulenta", MC_0) ~ "fraudulenta",
    #                        grepl("pungens", MC_0) ~ "pungens",
   #                         grepl("australis", MC_0) ~ "australis",
     #                       grepl("calliantha", MC_0) ~ "calliantha")) #%>% 
  #select(PR2, MC)

library(kableExtra)

cfr_taxonomy_shared %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

The two taxonomies do not correspond. Why?

<br>


5.2. Let's look at the overall taxonomic repartition in the two datasets.

<div class="column-left">
```{r warning=FALSE, message=FALSE, cache=TRUE}
library(treemapify)

final_tara_pr2 %>% 
  group_by(species) %>% 
  summarise(richness = n_distinct(seqId)) %>% 
  ungroup() %>%
  ggplot(aes(area=richness, fill=species, label=species)) +
  geom_treemap(show.legend = F) + 
 geom_treemap_text(fontface = "italic",
                    color = "black") +
                   # place = "center") +
  scale_fill_brewer(palette = "Blues") +
  ggtitle(paste0("Taxonomical ripartition of the ",
                 n_distinct(final_tara_pr2$seqId), " amplicons (PR2 ref.)"),
          paste0(n_distinct(final_tara_pr2$Station.label), " sampling stations;",
                 sum(final_tara_pr2$readcounts), "reads"))


#final_tara_pr2 %>% 
 # ggplot() +
 # aes(x = "", y = tot.numb.reads, fill = species) +
 # geom_bar(width = 1, stat = "identity") +
 # coord_polar("y", start=0) +
 # theme_bw() +
 # scale_fill_brewer(palette = "Paired") +
 
```
</div><div class="column-right">
```{r warning=FALSE, message=FALSE, cache=TRUE}
#tassonomia mia:
library(RColorBrewer)

nb.cols <- 13
mycolors <- colorRampPalette(brewer.pal(9,"BuGn"))(nb.cols)

final_tara_mc %>%
  group_by(species) %>% 
  summarise(richness = n_distinct(seqId),
            occupancy = n_distinct(Station.label)) %>%
  ungroup() %>%
  mutate(species_sel = ifelse(occupancy < 50, "others (*)",
                              species)) %>% 
  ungroup() %>%
  group_by(species_sel) %>% 
  summarise(richness2 = sum(richness)) %>% 
  ggplot(aes(area=richness2, fill=species_sel, label=species_sel)) +
  geom_treemap(show.legend = F) + 
  geom_treemap_text(fontface = "italic",
                    color = "black") +
  # place = "center") +
  #scale_fill_brewer(palette = "BuGn") +
  scale_fill_manual(values = mycolors) +
  ggtitle(paste0("Taxonomical ripartition of the ",
                 n_distinct(final_tara_mc$seqId), " amplicons (MC ref.)"),
          paste0(n_distinct(final_tara_mc$Station.label), " sampling stations; MC annotation; ",
                 sum(final_tara_mc$readcounts), "reads")) +
  labs(caption = "(*) species occurring in less than 50 stations")

```
</div>

```{r eval=FALSE, message=FALSE, cache=TRUE}
print("ue")
```



<br>

5.3. Let's look at the **geographical distribution** of the species
```{r warning=FALSE, message=FALSE, cache=TRUE, fig.height=10, fig.width=12}
###devo unirli in un unico file e facewrappare tutto:


library(scatterpie)

combined_datasets_mappie <- rbind(final_tara_pr2, final_tara_mc) %>% 
  group_by(Station.label, dataset) %>% 
  mutate(tot.abundance.per.station = sum(readcounts),
         Lat = mean(Latitude),
         Long = mean(Longitude)) %>% 
  ungroup() %>% 
  group_by(species, dataset) %>% 
  mutate(occupancy = n_distinct(Station.label)) %>% 
  ungroup() %>% 
  mutate(species_sel = ifelse(occupancy < 50, "others (*)",
                              as.character(species))) %>% 
  group_by(species_sel, Station.label, dataset) %>% 
  mutate(sel.species.abundance.per.station = sum(readcounts),
         fraction = sel.species.abundance.per.station/tot.abundance.per.station) %>% 
  ungroup() %>% 
  select(Station.label, Long, Lat, species_sel, fraction, dataset) %>% 
  arrange(Station.label) %>% 
  unique()  %>% 
  group_by(Station.label) %>% 
  mutate(radius = max(fraction)*4) %>% 
  ungroup() %>% 
  spread(species_sel, fraction) %>% 
  mutate_all(funs(ifelse(is.na(.), 0, .)))


mypalette <- randomcoloR::distinctColorPalette(17)

base_world3 + 
  geom_scatterpie(aes(x=Long, y=Lat, r=radius),
                  data=combined_datasets_mappie,
                 # cols= c("arenysensis", "calliantha", "cuspidata", "pungens",
                #          "seriata", "unknown", "americana",
                #          "australis", "micropora", "delicatissima",
                 #         "galaxiae", "brasiliana", "others (*)"),
                  #alpha=.7) +
                cols = c("calliantha", "cuspidata", "decipiens","delicatissima",
                         "fraudulenta", "galaxiae", "ID1_aren|dolo|arc|subcur",
                         "ID2_amer|cacia|circump|granii", "ID3_australis|multiseries", 
                         "ID4_brasiliana|lineola", "micropora", "others (*)", 
                         "pungens", "sabit", "seriata", "turgidula", "unknown"),
                         alpha = .7) +
  theme_minimal() +
  theme(legend.text=element_text(size=10)) +
  coord_fixed() +
 # scale_fill_brewer(palette = "Paired") +
  scale_fill_manual(values = mypalette) +
  facet_wrap(~dataset, nrow = 2) +
  ggtitle("Pseudo-nitzschia abundance and distribution at species level")  +
  labs(caption = "(*) species occurring in less than 50 stations")
#+
  #geom_scatterpie_legend(combined_datasets_mappie$radius, x=-150, y=-70)
```


<br>


6. What about **size classes**? Which one holds the highest number of sequences? Is there overlap among different classes?


```{r warning=FALSE, message=FALSE, cache=TRUE}

print("Combining the two dataset; Overlap considering seqId")

rbind(final_tara_pr2, final_tara_mc) %>% 
  unite("Fraction.size", lower.size.fraction:upper.size.fraction, sep = "-") %>% 
  mutate(Fraction.size = case_when(Fraction.size %in% c("3-20", "5-20") ~ "nano",
                                 Fraction.size == "20-180" ~ "micro",
                                 Fraction.size == "180-2000" ~ "meso",
                                 TRUE ~ "other")) %>%
  select(seqId, readcounts, Fraction.size) %>%
  group_by(seqId, Fraction.size) %>%
  summarise(sum = sum(readcounts)) %>%
  ungroup() %>%
  mutate(sum = ifelse(sum > 0,  1, 0)) %>%
  spread(Fraction.size, sum) %>%
  mutate_at(vars(-seqId), funs(ifelse(is.na(.), 0, .))) %>%
  select(-c("seqId", "other")) %>%
  venn::venn(ellipse = F, borders = F,  cexil = 0.8, cexsn = 0.9)
```


## Extras

##### Occurrence of sequences

```{r warning=FALSE, message=FALSE, cache=TRUE}
rbind(final_tara_pr2, final_tara_mc) %>% 
  mutate(Station.label = substr(Station.label, 6, 8)) %>% 
  group_by(seqId, dataset) %>% 
  summarise(occurrence = n_distinct(Station.label)) %>% 
  ggplot() +
  geom_histogram(aes(occurrence, fill=dataset), alpha=0.5) +
  theme(legend.position = "none") +
  theme_bw() +
  scale_x_log10() +
  xlab("Number of stations") +
  scale_fill_manual(values = c("TARA-MC" = "lightgreen",
                               "TARA-PR2" = "skyblue")) 


rbind(final_tara_pr2, final_tara_mc) %>% 
  mutate(Station.label = substr(Station.label, 6, 8)) %>% 
  group_by(seqId, dataset) %>% 
  summarise(occurrence = n_distinct(Station.label)) %>% 
  filter(occurrence <2) 


```

There are 15 sequences occurring only in one station.

<br>

<br>


##### **Explore PR2 Database**

```{r warning=FALSE, message=FALSE, cache=TRUE, fig.align='center'}

library(data.table)
library(tidyverse)

pr2_pn <- fread("C:/Users/Userszn/Documents/PhD/script/Pseudonitzschia/metaB/V4/pr2_version_4.11.1_UniEuk_V4_unique_Pn.tsv", header  = FALSE)

colnames(pr2_pn) <- c("lineage", "sequence", "taxogroup1", "taxogroup2", "PR2_id", "GenBank_ID")

pr2_pn_final <- splitstackshape::cSplit(pr2_pn, "lineage", "|") %>% 
  mutate(species = ifelse(is.na(lineage_13), "Pseudo-nitzschia+unknown",
                          as.character(lineage_13))) %>% 
  select(-lineage_13) %>% 
  splitstackshape::cSplit("species", "+") %>%
  rename(species = species_2) %>% 
  mutate(seq_length = nchar(sequence)) %>% 
  select(sequence, seq_length, species, PR2_id, GenBank_ID)

paste0(n_distinct(pr2_pn_final$sequence), " identical sequences")
paste0(n_distinct(pr2_pn_final$species), " species detected")
paste0(n_distinct(pr2_pn_final$PR2_id), " PR2 IDs")


pr2_pn_final1<- splitstackshape::cSplit(pr2_pn_final, "GenBank_ID", ",")

genbank_abb <- as.data.frame(lapply(pr2_pn_final1[,5:9], function(y) gsub("\\..*", "", y)))

pr2_pn_final2 <- pr2_pn_final1 %>% 
  select(sequence:PR2_id) %>% 
  cbind(genbank_abb) 


pr2_pn_final3 <- pr2_pn_final2 %>% 
  pivot_longer(GenBank_ID_1:GenBank_ID_5) %>% 
  #mutate(ID = ifelse(grepl("GenBank", name), "GenBank_ID", "altro")) %>% 
  select(-name) %>% 
  rename(GenBank_ID = value) %>% 
  na.omit()

write.csv(pr2_pn_final3, "C:/Users/Userszn/Documents/PhD/script/Pseudonitzschia/metaB/V4/pr2_pn_final3.csv", row.names = F)

pr2_pn_final_genbankinfo <- read.csv("C:/Users/Userszn/Documents/PhD/script/Pseudonitzschia/metaB/V4/pr2_pn_final_genbankinfo.csv") %>% na.omit()

pr2_pn_final_genbankinfo %>% 
  select(species, PR2_id, GenBank_ID_input, Origin) %>% 
  mutate(colture = ifelse(Origin == "cultured", 1, 0)) %>% 
  group_by(species) %>% 
  summarise(PR2 = n_distinct(PR2_id),
            GenBank = n_distinct(GenBank_ID_input),
            GenBank_cultured = sum(colture)) %>% 
  ungroup() %>% 
  mutate(GenBank_uncultured = GenBank - GenBank_cultured) %>% 
  select(-GenBank) %>% 
  pivot_longer(PR2:GenBank_uncultured) %>% 
  rename(ID = name) %>% 
  ggplot() +
  aes(x=value, y=reorder(species, desc(species)), fill=ID) +
  geom_bar(stat="identity", position = "dodge") +
  theme_bw() +
  xlab("") +
  ylab("") +
  #scale_fill_manual(values = c("PR2" = "tomato",
   #                            "GenBank_cultured" = "forestgreen",
  #                             "GenBank_uncultured" = "olivedrab3")) +
  scale_x_continuous(breaks=c(1, 3, 5, 7, 9, 11 )) +
  ggtitle(paste0(n_distinct(pr2_pn_final2$PR2_id), " Reference Ids (PR2)"))

```
