---
title: "MareChiara microscopy exploration "
subtitle: " "
author: "Lucia Campese"
date: "April 2021"
output:
  html_document:
    code_folding: hide
---

  ```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<style>
.column-left{
  float: left;
  width: 50%;
  text-align: left;
}
.column-right{
  float: right;
  width: 50%;
  text-align: right;
}
</style>
  
<br>

### Aim
Explore microscopy counts of phytoplankton taxa from the LTER MareChiara

### Dataset
##### Data: LTER MareChiara microscopy data


### Workflow

<br>


1. Upload datasets

```{r warning=FALSE, message=FALSE, cache=TRUE, fig.align='center'}
#packages
library(tidyverse)
library(tidyquant)
library(timetk)

#upload files
MC_species_counts <- readxl::read_xlsx("C:/Users/Userszn/Documents/PhD/script/Pseudonitzschia/MC_microscopy/Pn_and_selected_species_1984_2019.xlsx",
                                       skip = 1, col_names = T)

sampling_dates <- readxl::read_xlsx("C:/Users/Userszn/Documents/PhD/script/Pseudonitzschia/MC_microscopy/SamplingDates_1984_2019.xlsx",
                                    col_names = FALSE, col_types = "date")

sampling_dates_2 <- sampling_dates %>% 
  rename(Date = ...1)

sample_IDs <- MC_species_counts %>%
  rename(tax = Sample_ID) %>%
  pivot_longer(2:1323) %>%
  splitstackshape::cSplit("name", sep = " ") %>% 
  mutate(name_3 = ifelse(is.na(name_3), name_2, name_3),
         name_4 = sprintf("%04d", as.numeric(name_2)),
         name_5 = ifelse(name_2 == name_3, name_4, paste0(name_4, "_", name_3)),
         Sample_ID = paste0(name_1, "_", name_5)) %>% 
  select(Sample_ID) %>% 
  unique()

sample_IDS_dates <- cbind(sample_IDs, sampling_dates_2)

MC_species_counts_long <- MC_species_counts %>%
  rename(tax = Sample_ID) %>%
  pivot_longer(2:1323) %>%
  splitstackshape::cSplit("name", sep = " ") %>% 
  mutate(name_3 = ifelse(is.na(name_3), name_2, name_3),
         name_4 = sprintf("%04d", as.numeric(name_2)),
         name_5 = ifelse(name_2 == name_3, name_4, paste0(name_4, "_", name_3)),
         Sample_ID = paste0(name_1, "_", name_5)) %>% 
  select(tax, Sample_ID, value) %>% 
  splitstackshape::cSplit("tax", sep = " ") %>% 
  rename(genus = tax_1) %>% 
  mutate(species = case_when(tax_2 == "cf." ~ "cf.pungens",
                             tax_2 == "danicus" ~ "danicus_complex",
                             tax_2 == "delicatissima" ~ "delicatissima_complex",
                             (tax_2 == "galaxiae" & grepl("small", tax_3)) ~ "galaxiae_small",
                             tax_2 == "pseudodlicatissima" ~ "pseudodelicatissima_complex",
                             TRUE ~ as.character(tax_2))) %>% 
  select(-c(tax_2:tax_4)) %>% 
  filter(value > 0)


MC_species_counts_long_2 <- MC_species_counts_long %>% 
  left_join(sample_IDS_dates, by = "Sample_ID")


#remove Leoptocylindrus aporus e Pseudonitzschia cf pungens
#make tax column
#save output 
#MC_species_counts_long_3 <- MC_species_counts_long_2 %>%
  #unite("tax", genus:species, sep = "_") %>%
  #filter(!tax %in% c("Leptocylindrus_aporus", "Pseudo-nitzschia_cf.pungens"))

```

2. Retrieve taxonomy information per each species

```{r, eval=FALSE}

#tidy taxonomy
MC_species_taxonomy_worms <-  MC_species_counts_long_2 %>% 
  mutate(genus = as.character(genus)) %>% 
  select(genus) %>% 
  pull() %>% 
  unique() %>% 
  enframe(name = NULL) %>% 
  mutate(taxo = map(value, taxize::classification, db = "worms")) %>% 
  mutate(taxa = map(taxo, function(x) x[[1]])) %>% 
  unnest(taxa) %>% 
  select(value, name, rank)

save(MC_species_taxonomy_worms, file = "C:/Users/Userszn/Documents/PhD/script/Pseudonitzschia/MC_microscopy/MC_species_taxonomy_worms.RData")
```


```{r warning=FALSE, message=FALSE, cache=TRUE, fig.align='center'}
load("C:/Users/Userszn/Documents/PhD/script/Pseudonitzschia/MC_microscopy/MC_species_taxonomy_worms.RData")

MC_species_taxonomy_worms_wide <-  MC_species_taxonomy_worms %>% 
  pivot_wider(names_from = rank, values_from = name) %>% 
  mutate_at(vars(4:15), ~ifelse(is.na(.), "unknown", .)) %>% 
  mutate(Phylum = ifelse(Phylum == "unknown", "Chlorophyta", Phylum)) %>% 
  select(Kingdom, Phylum, Class, Order, Family, Genus)

MC_species_counts_long_2_tax <- MC_species_counts_long_2 %>% 
  rename(Genus = genus) %>% 
  inner_join(MC_species_taxonomy_worms_wide, by = "Genus") %>% 
  select(Kingdom:Family, Genus, species, Sample_ID, Date, value)



print(paste0("The dataset comprises ",
      n_distinct(MC_species_counts_long_2_tax$species), " species belonging to ",
      n_distinct(MC_species_counts_long_2_tax$Genus), " genera ascribed to ",
      n_distinct(MC_species_counts_long_2_tax$Phylum), " phyla: ",
      unique(MC_species_counts_long_2_tax$Phylum)[1], ", ",
      unique(MC_species_counts_long_2_tax$Phylum)[2], ", ",
      unique(MC_species_counts_long_2_tax$Phylum)[3], ", ",
      unique(MC_species_counts_long_2_tax$Phylum)[4], ", ",
      unique(MC_species_counts_long_2_tax$Phylum)[5]))


```


3. Check the occurrence of each taxon in each sample

### Genus level {.tabset}

#### Class: Bacillariophyceae 
```{r warning=FALSE, message=FALSE, cache=TRUE, fig.align='center'}

#Occurrence
#per genus (col by Family)
#Bacillariophyceae
MC_species_counts_long_2_tax %>%
  mutate(genusab = substr(Genus, 1, 1)) %>% 
  filter(Class == "Bacillariophyceae") %>% 
  unite("tax", c(genusab,species), sep = ".") %>% 
  group_by(Phylum, Class, Order, Family, Genus) %>% 
  summarise(n_samples = n_distinct(Sample_ID)) %>% 
  ungroup() %>% 
  ggplot() +
  aes(x=Genus, y=n_samples, fill = Family) +
  geom_col() +
  theme_bw() +
  theme(axis.text.x = element_text(angle=60, size=8, vjust = 0.5)) +
  xlab("Genus") +
  ylab("Occurrence") + 
  scale_fill_brewer(palette = "Paired") +
  facet_wrap(~ paste0("Phylum: ", Phylum, "; Class: " , Class))

```

#### Class: Prymnesiophyceae

```{r warning=FALSE, message=FALSE, cache=TRUE, fig.align='center'}
#Prymnesiophyceae
MC_species_counts_long_2_tax %>%
  mutate(genusab = substr(Genus, 1, 1)) %>% 
  filter(Class == "Prymnesiophyceae") %>% 
  unite("tax", c(genusab,species), sep = ".") %>% 
  group_by(Phylum, Class, Order, Family, Genus) %>% 
  summarise(n_samples = n_distinct(Sample_ID)) %>% 
  ungroup() %>% 
  ggplot() +
  aes(x=Genus, y=n_samples, fill = Family) +
  geom_col() +
  theme_bw() +
  theme(axis.text.x = element_text(angle=60, size=8, vjust = 0.5)) +
  xlab("Genus") +
  ylab("Occurrence") + 
  scale_fill_brewer(palette = "Paired") +
  facet_wrap(~ paste0("Phylum: ", Phylum, "; Class: " , Class))

```

#### Other Classes
```{r warning=FALSE, message=FALSE, cache=TRUE, fig.align='center', fig.width=8}
#Other 
MC_species_counts_long_2_tax %>%
  mutate(genusab = substr(Genus, 1, 1)) %>% 
  filter(!Class %in% c("Bacillariophyceae", "Prymnesiophyceae")) %>% 
  unite("tax", c(genusab,species), sep = ".") %>% 
  group_by(Phylum, Class, Order, Family, Genus) %>% 
  summarise(n_samples = n_distinct(Sample_ID)) %>% 
  ungroup() %>% 
  ggplot() +
  aes(x=Genus, y=n_samples, fill = Family) +
  geom_col() +
  theme_bw() +
  theme(axis.text.x = element_text(angle=60, size=8, vjust = 0.5)) +
  xlab("Genus") +
  ylab("Occurrence") + 
  scale_fill_brewer(palette = "Paired") +
  facet_wrap(~ paste0("Phylum: ", Phylum, "; Class: " , Class), ncol = 2)
```


### Species level {.tabset}

#### Class: Bacillariophyceae 

```{r warning=FALSE, message=FALSE, cache=TRUE, fig.align='center', fig.width=8}
#per species (col by Genus)
#Bacillariophyceae
MC_species_counts_long_2_tax %>%
  mutate(genusab = substr(Genus, 1, 1)) %>% 
  filter(Class == "Bacillariophyceae") %>% 
  unite("tax", c(genusab,species), sep = ".") %>% 
  group_by(Phylum, Class, Order, Family, Genus, tax) %>% 
  summarise(n_samples = n_distinct(Sample_ID)) %>% 
  ungroup() %>% 
  ggplot() +
  aes(x=tax, y=n_samples, fill = Genus) +
  geom_col() +
  theme_bw() +
  theme(axis.text.x = element_text(angle=60, size=8, vjust = 0.5)) +
  xlab("Species") +
  ylab("Occurrence") + 
  scale_fill_brewer(palette = "Paired") +
  facet_wrap(~ paste0("Phylum: ", Phylum, "; Class: " , Class))
```

#### Class: Prymnesiophyceae
```{r warning=FALSE, message=FALSE, cache=TRUE, fig.align='center'}
#Prymnesiophyceae
MC_species_counts_long_2_tax %>%
  mutate(genusab = substr(Genus, 1, 1)) %>% 
  filter(Class == "Prymnesiophyceae") %>% 
  unite("tax", c(genusab,species), sep = ".") %>% 
  group_by(Phylum, Class, Order, Family, Genus, tax) %>% 
  summarise(n_samples = n_distinct(Sample_ID)) %>% 
  ungroup() %>% 
  ggplot() +
  aes(x=tax, y=n_samples, fill = Genus) +
  geom_col() +
  theme_bw() +
  theme(axis.text.x = element_text(angle=60, size=8, vjust = 0.5)) +
  xlab("Species") +
  ylab("Occurrence") + 
  scale_fill_brewer(palette = "Paired") +
  facet_wrap(~ paste0("Phylum: ", Phylum, "; Class: " , Class))
```

#### Other Classes

```{r warning=FALSE, message=FALSE, cache=TRUE, fig.align='center'}
#Others
MC_species_counts_long_2_tax %>%
  mutate(genusab = substr(Genus, 1, 1)) %>% 
  filter(!Class %in% c("Bacillariophyceae", "Prymnesiophyceae")) %>% 
  unite("tax", c(genusab,species), sep = ".") %>% 
  group_by(Phylum, Class, Order, Family, Genus, tax) %>% 
  summarise(n_samples = n_distinct(Sample_ID)) %>% 
  ungroup() %>% 
  ggplot() +
  aes(x=tax, y=n_samples, fill = Genus) +
  geom_col() +
  theme_bw() +
  theme(axis.text.x = element_text(angle=60, size=8, vjust = 0.5)) +
  xlab("Species") +
  ylab("Occurrence") + 
  scale_fill_brewer(palette = "Paired") +
  facet_wrap(~ paste0("Phylum: ", Phylum, "; Class: " , Class), ncol = 2)
```

###
```{r warning=FALSE, message=FALSE, cache=TRUE, fig.align='center'}

#remove Leoptocylindrus aporus e Pseudonitzschia cf pungens
#make tax column
#save output 
MC_species_counts_long_3_tax <- MC_species_counts_long_2_tax %>%
  unite("tax", Genus:species, sep = "_") %>%
  filter(!tax %in% c("Leptocylindrus_aporus", "Pseudo-nitzschia_cf.pungens"))
```


4. Explore the distribution of each species[*] through time

[*]Note: I removed L.aporus and P.cf.pungens (too low occurrence)


### Time-series analysis{.tabset}
#### Class: Bacillariophyceae 

```{r warning=FALSE, message=FALSE, cache=TRUE, fig.align='center'}
#1. Investigate a time series
#make normal time series plot
#each taxon
Bacillariophyceae_timeseriesplot <- MC_species_counts_long_3_tax %>%
  filter(Class == "Bacillariophyceae") %>% 
  group_by(tax) %>% 
  nest() %>% 
  mutate(plot = map2(tax, data, ~ plot_time_series(.data = .y,
                                                   Date, 
                                                   value, 
                                                   .interactive = FALSE,
                                                   .smooth = FALSE) +
                       xlab("Date") +
                       ylab("cells/ml") +
                       ggtitle(.x)))


ggsave(file="C:/Users/Userszn/Documents/PhD/script/Pseudonitzschia/MC_microscopy/Bacillariophyceae_timeseriesplot.pdf",
       width = 12,
       height = 12,
       gridExtra::marrangeGrob(grobs = Bacillariophyceae_timeseriesplot$plot,
                               nrow=2, ncol=2,
                               top = "Bacillariophyceae: Time series plot"))

```

![timeseriesplot](C:/Users/Userszn/Documents/PhD/script/Pseudonitzschia/MC_microscopy/Bacillariophyceae_timeseriesplot.pdf){width=100% height=700}

#### Class: Prymnesiophyceae 

```{r warning=FALSE, message=FALSE, cache=TRUE, fig.align='center'}
#1. Investigate a time series
#make normal time series plot
#each taxon
Prymnesiophyceae_timeseriesplot <- MC_species_counts_long_3_tax %>%
  filter(Class == "Prymnesiophyceae") %>% 
  group_by(tax) %>% 
  nest() %>% 
  mutate(plot = map2(tax, data, ~ plot_time_series(.data = .y,
                                                   Date, 
                                                   value, 
                                                   .interactive = FALSE,
                                                   .smooth = FALSE) +
                       xlab("Date") +
                       ylab("cells/ml") +
                       ggtitle(.x)))


ggsave(file="C:/Users/Userszn/Documents/PhD/script/Pseudonitzschia/MC_microscopy/Prymnesiophyceae_timeseriesplot.pdf",
       width = 12,
       height = 12,
       gridExtra::marrangeGrob(grobs = Prymnesiophyceae_timeseriesplot$plot,
                               nrow=2, ncol=2,
                               top = "Prymnesiophyceae: Time series plot"))

```

![timeseriesplot](C:/Users/Userszn/Documents/PhD/script/Pseudonitzschia/MC_microscopy/Prymnesiophyceae_timeseriesplot.pdf){width=100% height=700}




### Seasonality plots{.tabset}

#### Class: Bacillariophyceae 

```{r warning=FALSE, message=FALSE, cache=TRUE, fig.align='center'}
#3. Make a seasonality plot
#seasonal diagnostics
#each taxon
Bacillariophyceae_seasonal_diagnositcs <- MC_species_counts_long_3_tax %>%
  filter(Class == "Bacillariophyceae") %>% 
  group_by(tax) %>% 
  nest() %>%  
  mutate(plot = map2(tax, data, ~ plot_seasonal_diagnostics(.data = .y,
                                                            Date, value, .interactive = FALSE,
                                                            .feature_set = c("week", "month.lbl", "quarter", "year"),
                                                            .geom = "boxplot",
                                                            .y_lab = "cells/ml") +
                       theme(axis.text.x = element_text(angle=50, size=9, vjust = 0.5)) +
                       ggtitle(.x))) 


ggsave(file="C:/Users/Userszn/Documents/PhD/script/Pseudonitzschia/MC_microscopy/Bacillariophyceae_seasonal_diagnositcs.pdf",
       width = 9,
       height = 12,
       gridExtra::marrangeGrob(grobs = Bacillariophyceae_seasonal_diagnositcs$plot,
                               nrow=1, ncol=1,
                               top = "Bacillariophyceae: Seasonal diagnostics"))

```
![Bacillariophyceae_seasonal_diagnositcs](C:/Users/Userszn/Documents/PhD/script/Pseudonitzschia/MC_microscopy/Bacillariophyceae_seasonal_diagnositcs.pdf){width=100% height=700}


#### Class: Prymnesiophyceae 

```{r warning=FALSE, message=FALSE, cache=TRUE, fig.align='center'}
#3. Make a seasonality plot
#seasonal diagnostics
#each taxon
Prymnesiophyceae_seasonal_diagnositcs <- MC_species_counts_long_3_tax %>%
  filter(Class == "Prymnesiophyceae") %>% 
  group_by(tax) %>% 
  nest() %>%  
  mutate(plot = map2(tax, data, ~ plot_seasonal_diagnostics(.data = .y,
                                                            Date, value, .interactive = FALSE,
                                                            .feature_set = c("week", "month.lbl", "quarter", "year"),
                                                            .geom = "boxplot",
                                                            .y_lab = "cells/ml") +
                       theme(axis.text.x = element_text(angle=50, size=9, vjust = 0.5)) +
                       ggtitle(.x))) 


ggsave(file="C:/Users/Userszn/Documents/PhD/script/Pseudonitzschia/MC_microscopy/Prymnesiophyceae_seasonal_diagnositcs.pdf",
       width = 9,
       height = 12,
       gridExtra::marrangeGrob(grobs = Prymnesiophyceae_seasonal_diagnositcs$plot,
                               nrow=1, ncol=1,
                               top = "Prymnesiophyceae: Seasonal diagnostics"))

```
![Bacillariophyceae_seasonal_diagnositcs](C:/Users/Userszn/Documents/PhD/script/Pseudonitzschia/MC_microscopy/Prymnesiophyceae_seasonal_diagnositcs.pdf){width=100% height=700}

