---
title: "*Pseudonitzschia* and its parasitoids"
subtitle: "Global scale co-occurrence analysis"
author: "Lucia Campese"
date: "September 2020"
output:
  html_document:
    code_folding: hide
---
---
  
  ```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<style>
.column-left{
  float: left;
  width: 50%;
  text-align: left;
}
.column-right{
  float: right;
  width: 50%;
  text-align: right;
}
</style>
  
<br>


### Aim
The aim is to explore the *in silico* association between *Pseudonitzschia* and its known eukaryotic parasitoids through a host-parasite network using Tara Oceans data.

<br>

### Dataset
##### Data: TARA Oceans and TARA Oceans Polar circle metabarcording (V4)
##### Taxa: *Pseudonitzschia*, Fungi, Oomycetes
##### Geographic scale: Global


<br>

#### **Preparing datasets**

<br>

##### 1. *Pseudonitzschia*

The used dataset is made of **324** different **sequences** assigned to *Pseudonitzschia* species with a **100% identity** to MareChiara database (curated by Valeria Ruggiero) and occurring in at least **3 sampling stations** in the TARA Oceans dataset (see [here](https://htmlpreview.github.io/?https://github.com/Iucia/Pseudonitzschia/blob/master/comparison100identity.html "Pn_MC-TARA") for details on dataset building).

```{r warning=FALSE, message=FALSE, cache=TRUE}

library(data.table)
library(tidyverse)

Pn_candi_identity100.distr <- fread("C:/Users/Userszn/Documents/PhD/script/Pseudonitzschia/metaB/V4/cfr100pid/18S_V4_1191_samples_Pncandi_3readsmin_pid100.distr") 

colnames(Pn_candi_identity100.distr) <- c("seqId", "sample", "readcounts")

paste0(n_distinct(Pn_candi_identity100.distr$seqId), " Pseudonitzschia sequence Ids")

Pn_candi_identity100.tax <- fread("C:/Users/Userszn/Documents/PhD/script/Pseudonitzschia/metaB/V4/cfr100pid/Pn_tara_candi_final_pid100.txt") %>%
  separate(V1, into = c("seqId", "size"), sep = ";") %>% 
  rename(tax = V2, pid = V3) %>% 
  select(seqId, tax, pid)

#subset tax (still with 508 seq) using the distr file (with the corret 324 seq) {the aim is to add the taxonomy info}

Pn_candi_identity100.distr.tax <- inner_join(Pn_candi_identity100.tax, Pn_candi_identity100.distr, by = "seqId")

##add env

env <- fread("C:/Users/Userszn/Documents/PhD/script/Pseudonitzschia/metaB/V4/18S_V4_1191_samples_context.tsv")

Pn_candi_identity100.distr.tax.env <- inner_join(Pn_candi_identity100.distr.tax, env) %>% 
  select(seqId:readcounts,Pangaea_id:minimum.generation.time.h)


final_tara_mc <- Pn_candi_identity100.distr.tax.env %>% 
  mutate(species = case_when(grepl("allochrona", tax) ~ "allochrona",
                             grepl("americana", tax) ~ "ID2_amer|cacia|circump|granii",
                             grepl("arenysensis", tax) ~ "ID1_aren|dolo|arc|subcur",
                             grepl("ID3_P._australis", tax) ~ "ID3_australis|multiseries",
                             grepl("P._australis_SPC21", tax) ~ "australis_single",
                             grepl("batesiana", tax) ~ "batesiana",
                             grepl("brasiliana", tax) ~ "ID4_brasiliana|lineola",
                             grepl("calliantha", tax) ~ "calliantha",
                             grepl("cuspidata", tax) ~ "cuspidata",
                             grepl("decipiens", tax) ~ "decipiens",
                             grepl("delicatissima", tax) ~ "delicatissima",
                             grepl("fraudulenta", tax) ~ "fraudulenta",
                             grepl("fukuyoi", tax) ~ "fukuyoi",
                             grepl("galaxiae", tax) ~ "galaxiae",
                             grepl("heimii", tax) ~ "heimii",
                             grepl("kodamae", tax) ~ "ID5_kodamae|lundholmiae",
                             grepl("mannii", tax) ~ "mannii",
                             grepl("micropora", tax) ~ "micropora",
                             grepl("P._multiseries_NparI", tax) ~ "multiseries_single",
                             grepl("multistriata", tax) ~ "multistriata",
                             grepl("pseudodelicatissima", tax) ~ "pseudodelicatissima",
                             grepl("pungens", tax) ~ "pungens",
                             grepl("sabit", tax) ~ "sabit",
                             grepl("turgidula", tax) ~ "turgidula",
                            TRUE ~ "ERRORE")) %>% 
  mutate(length = "tobedone") %>% 
  select(seqId, length, species, sample, readcounts, Station.label:Depth.chloro.max) %>%
  mutate(dataset = c("TARA-MC"))
```

<br>

<br>


<br>

##### 2. Oomycetes

The initial dataset is made of **2997** different **sequences** assigned to Peronosporomycetes with a variable percentage of identity to PR2 database and occurring in at least **3 sampling stations** in the TARA Ocean dataset. 

```{bash, eval = FALSE}
#1. Extract Peronosporomycetes  sequences from the the V4

srun cat 18S_V4_1191_samples.results | grep Peronosporomycetes  > 18S_V4_1191_samples_oomyc.results &
  
  
#take only sequences with at least 3 reads (seen three times in the whole dataset) [*]
srun awk '$2>2' 18S_V4_1191_samples_oomyc.results > 18S_V4_1191_samples_oomyc_3readsmin.results &
  
#wc -l 5403 18S_V4_1191_samples_oomyc.results
#wc -l 4451 18S_V4_1191_samples_oomyc_3readsmin.results
  
#[*]as adviced by Nicholas Henry.
  

#Extract only seq ids (column 1)
awk '{print $1}' 18S_V4_1191_samples_oomyc_3readsmin.results > 18S_V4_1191_samples_oomyc_3readsmin_seqid.txt

#wc -l: 4451 18S_V4_1191_samples_oomyc_3readsmin_seqid.txt


#use the retrieved 4451 sequence Ids to subset the distr file (i.e. the one 
#containing seqId|sampleId|readcounts

srun grep -Ff 18S_V4_1191_samples_oomyc_3readsmin_seqid.txt ~/Data/metaB/V4/18S_V4_1191_samples.distr > 18S_V4_1191_samples_oomyc_3readsmin.distr &
  

```

```{r warning=FALSE, message=FALSE, cache=TRUE}
library(data.table)
library(tidyverse)

oomyc.distr <- fread("C:/Users/Userszn/Documents/PhD/script/Pseudonitzschia/metaB/V4/Parasites/Oomiceti/18S_V4_1191_samples_oomyc_3readsmin.distr")


oomyc.tax <- fread("C:/Users/Userszn/Documents/PhD/script/Pseudonitzschia/metaB/V4/Parasites/Oomiceti/18S_V4_1191_samples_oomyc_3readsmin.results", header = F, sep = "\t") #17 sequences


colnames(oomyc.distr) <- c("seqId", "sample", "readcounts")
colnames(oomyc.tax) <- c("seqId", "totreads", "perc.identity", "lineage", "idpr2")

#oomyc.tax %>% 
  #splitstackshape::cSplit("lineage", "|")

oomyc.distr.tax <- inner_join(oomyc.distr, oomyc.tax, by = "seqId") %>% 
  select(-totreads)

env <- fread("C:/Users/Userszn/Documents/PhD/script/Pseudonitzschia/metaB/V4/18S_V4_1191_samples_context.tsv")

oomyc.distr.tax.env <- inner_join(oomyc.distr.tax, env)


##filter on occurrence

oomyc.distr.tax.env.filtered <- oomyc.distr.tax.env %>% 
  group_by(seqId) %>% 
  mutate(Occurrence = n_distinct(Station.label)) %>% 
  ungroup() %>% 
  filter(Occurrence > 1)

oomyc.distr.tax.env1 <- oomyc.distr.tax.env.filtered %>% 
  select(seqId, lineage, perc.identity, sample, readcounts, Station.label:Depth.chloro.max)


#oomyc.distr.tax.env1 is the final file with seqIds, taxonomical annotation, distribution and environmental info


paste0(n_distinct(oomyc.distr.tax.env1$seqId), " Oomycetes sequences")


oomyc_3reads2stationsmin_seqid <- oomyc.distr.tax.env1 %>% 
  select(seqId) %>% 
  unique()

write.csv(oomyc_3reads2stationsmin_seqid, "oomyc_3reads2stationsmin_seqid.csv", row.names = F)


###levo il nome della colonna su excel e esporto il file come txt file e lo porto su falkor 
```

```{r eval=FALSE}

###FALKOR
###take fastas

#wc -l 2997 oomyc_3reads2stationsmin_seqid.txt

##convert file (ascii with crlf terminators) into a simple ascii file, in order to be grepped with the fasta file (simple ascii):

cat oomyc_3reads2stationsmin_seqid.txt | tr -d '\r' > oomyc_3reads2stationsmin_seqid_ascii.txt

#use seqIds to retrieve fasta of the sequences

srun grep -w -A 1 -Ff oomyc_3reads2stationsmin_seqid_ascii.txt ~/Data/metaB/V4/18S_V4_1191_samples.fas --no-group-separator > 18S_V4_1191_samples_oomyc_3reads2stationsmin.fas &
  
#check if it retrieved the right number of sequences

grep ">" 18S_V4_1191_samples_oomyc_3reads2stationsmin.fas -c
#2997

wc -l oomyc_3reads2stationsmin_seqid_ascii.txt
#2997

#ok.

##I now have both the fasta: 18S_V4_1191_samples_oomyc_3reads2stationsmin.fas
##and the table with taxonomy, env.par etc: oomyc.distr.tax.env1
```

<div class="column-left">
```{r warning=FALSE, message=FALSE, cache=TRUE}
oomyc_percid <- oomyc.distr.tax.env1 %>% 
  select(seqId, perc.identity) %>% 
  unique() %>% 
  select(perc.identity) %>% 
  summary()


oomyc.distr.tax.env1 %>% 
 group_by(seqId) %>% 
  mutate(totreads = sum(readcounts)) %>% 
  ungroup() %>% 
  select(seqId, perc.identity, totreads) %>% 
  unique() %>% 
  ggplot + 
  aes(x = perc.identity, 
      fill = totreads) + 
  geom_histogram(bins = 100) + 
  theme_bw() + 
  annotation_custom(gridExtra::tableGrob(oomyc_percid), xmin=70, xmax=80, ymin=600, ymax=900) +
  ggtitle("Percentage of identity to the PR2 reference", 
          subtitle= paste0("Oomycetes, TARA V4; ", 
                           n_distinct(oomyc.distr.tax.env1$seqId), 
                           " distinct amplicons"))
```
</div><div class="column-right">
```{r warning=FALSE, message=FALSE, cache=TRUE}
oomyc.occurrences <- oomyc.distr.tax.env1 %>% 
  group_by(seqId) %>% 
  summarise(Occurrence = n_distinct(Station.label)) %>% 
  arrange(Occurrence) 

oomyc.table_occurrences <- oomyc.occurrences %>% 
  select(Occurrence) %>% 
  summary()

oomyc.occurrences %>% 
  ggplot() +
  aes(x=Occurrence) +
  geom_histogram(bins = 100) +
  theme_bw() +
  annotation_custom(gridExtra::tableGrob(oomyc.table_occurrences), xmin=25, xmax=55, ymin=400, ymax=750) +
  ggtitle("Occurrence of sequences along stations", 
          subtitle= paste0("Oomycetes, TARA V4; ", 
                           n_distinct(oomyc.occurrences$seqId), 
                           " distinct amplicons"))
```
</div>
```{r eval=FALSE}
library(maps)

world_map <- map_data("world")
p <- ggplot() + coord_fixed()

base_world3 <- p + geom_polygon(data=world_map,aes(x=long, y=lat,group=group),colour="gray80",size=0.2, fill="gray80")+
  theme_bw() + coord_fixed()


#map
oomyc.map <- oomyc.distr.tax.env1 %>% 
  group_by(Station.label) %>% 
  mutate(Lat= mean(Latitude),
         Long=mean(Longitude),
         abundance.per.station = sum(readcounts),
         richness.per.station.amplicons = n_distinct(seqId)) %>%
  ungroup() %>% 
  select(seqId, Station.label, Lat, Long, abundance.per.station,
         richness.per.station.amplicons) %>% unique()

base_world3+
  geom_point(data=oomyc.map,
             aes(x = Long, y = Lat, 
                 size = abundance.per.station,
                 fill = richness.per.station.amplicons),
             pch = 21, col = "black") +
  scale_fill_viridis_c() +
  theme_minimal() +
  theme(legend.text=element_text(size=10)) +
  ggtitle("Oomycetes abundance and richness per station (V4)", "") 


####map
fungi.map <- Fungi.distr.tax.env1 %>% 
  group_by(Station.label) %>% 
  mutate(Lat= mean(Latitude),
         Long=mean(Longitude),
         abundance.per.station = sum(readcounts),
         richness.per.station.amplicons = n_distinct(seqId)) %>%
  ungroup() %>% 
  select(seqId, Station.label, Lat, Long, abundance.per.station,
         richness.per.station.amplicons) %>% unique()

base_world3+
  geom_point(data=fungi.map,
             aes(x = Long, y = Lat, 
                 size = abundance.per.station,
                 fill = richness.per.station.amplicons),
             pch = 21, col = "black") +
  scale_fill_viridis_c() +
  theme_minimal() +
  theme(legend.text=element_text(size=10)) +
  ggtitle("Fungi abundance and richness per station (V4)", "") 






tbm_pseudonitzschia <- final_tara_mc %>% 
  select(seqId, sample:Depth.chloro.max) %>% 
  mutate(taxogroup = "Pseudonitzschia")

tbm_fungi <- Fungi.distr.tax.env1 %>% 
  select(seqId, sample:Depth.chloro.max) %>% 
  mutate(taxogroup = "Fungi")

tbm_oomyc <- oomyc.distr.tax.env1 %>% 
  select(seqId, sample:Depth.chloro.max) %>% 
  mutate(taxogroup = "Oomycetes") 



pn_oomyc_fungi.map <- rbind(tbm_pseudonitzschia, tbm_oomyc, tbm_fungi) %>% 
  group_by(Station.label) %>% 
  mutate(Lat= mean(Latitude),
         Long=mean(Longitude)) %>% 
  ungroup() %>% 
  group_by(Station.label, taxogroup) %>% 
  mutate(abundance.per.station = sum(readcounts),
         richness.per.station.amplicons = n_distinct(seqId)) %>%
  ungroup() %>% 
  select(seqId, Station.label, Lat, Long, abundance.per.station,
         richness.per.station.amplicons, taxogroup) %>% unique()

base_world3+
  geom_point(data=pn_oomyc_fungi.map,
             aes(x = Long, y = Lat, 
                 size = abundance.per.station,
                 fill = taxogroup),
             pch = 21, col = "black") +
  #scale_fill_viridis_c() +
  theme_minimal() +
  theme(legend.text=element_text(size=10)) +
  ggtitle("Abundance and richness per station (V4)", "") 
```

<br>

<br>



##### 3. Fungi

The initial dataset is made of **34523** different **sequences** assigned to Fungi with a variable percentage of identity to PR2 database and occurring in at least **3 sampling stations** in the TARA Oceans dataset. 
 

```{r warning=FALSE, message=FALSE, cache=TRUE}

library(data.table)
library(tidyverse)

Fungi.distr <- fread("C:/Users/Userszn/Documents/PhD/script/Pseudonitzschia/metaB/V4/Parasites/Fungi/18S_V4_1191_samples_Fungi_3readsmin.distr")

Fungi.tax <- fread("C:/Users/Userszn/Documents/PhD/script/Pseudonitzschia/metaB/V4/Parasites/Fungi/18S_V4_1191_samples_Fungi_3readsmin.results", header = F, sep = "\t") #17 sequences


colnames(Fungi.distr) <- c("seqId", "sample", "readcounts")
colnames(Fungi.tax) <- c("seqId", "totreads", "perc.identity", "lineage", "idpr2")


#Fungi.tax %>% 
#  splitstackshape::cSplit("lineage", "|") %>%
#  filter(lineage_06 != "Fungi")

##checked sono tutti funghi


Fungi.distr.tax <- inner_join(Fungi.distr, Fungi.tax, by = "seqId") %>%
  select(-totreads)


env <- fread("C:/Users/Userszn/Documents/PhD/script/Pseudonitzschia/metaB/V4/18S_V4_1191_samples_context.tsv")

Fungi.distr.tax.env <- inner_join(Fungi.distr.tax, env)


##filter on occurrence

Fungi.distr.tax.env.filtered <- Fungi.distr.tax.env %>% 
  group_by(seqId) %>% 
  mutate(Occurrence = n_distinct(Station.label)) %>% 
  ungroup() %>% 
  filter(Occurrence > 1)

Fungi.distr.tax.env1 <- Fungi.distr.tax.env.filtered %>% 
  select(seqId, lineage, perc.identity, sample, readcounts, Station.label:Depth.chloro.max)


#Fungi.distr.tax.env1 is the final file with seqIds, taxonomical annotation, distribution and environmental info


paste0(n_distinct(Fungi.distr.tax.env1$seqId), " Fungi sequences")


Fungi_3reads2stationsmin_seqid <- Fungi.distr.tax.env1 %>% 
  select(seqId) %>% 
  unique()

write.csv(Fungi_3reads2stationsmin_seqid, "Fungi_3reads2stationsmin_seqid.csv", row.names = F)


###levo il nome della colonna su excel e esporto il file come txt file e lo porto su falkor 

```

```{r, eval=FALSE}

###FALKOR
###take fastas

#wc -l 34523 Fungi_3reads2stationsmin_seqid.txt

##convert file (ascii with crlf terminators) into a simple ascii file, in order to be grepped with the fasta file (simple ascii):

cat Fungi_3reads2stationsmin_seqid.txt | tr -d '\r' > Fungi_3reads2stationsmin_seqid_ascii.txt

#use seqIds  to retrieve fasta of the sequences
srun grep -w -A 1 -Ff Symbionts/Fungi_3reads2stationsmin_seqid_ascii.txt 18S_V4_1191_samples.fas --no-group-separator > 18S_V4_1191_samples_Fungi_3reads2stationsmin.fas &
  
#check if it retrieved the right number of sequences

grep ">" 18S_V4_1191_samples_Fungi_3reads2stationsmin.fas -c
#34523
wc -l Fungi_3reads2stationsmin_seqid_ascii.txt
#34523

#ok.

##I now have both the fasta: 18S_V4_1191_samples_Fungi_3reads2stationsmin.fas
##and the table with taxonomy, env.par etc: Fungi.distr.tax.env1
```


<div class="column-left">
```{r warning=FALSE, message=FALSE, cache=TRUE}

Fungi_percid <- Fungi.distr.tax.env1 %>% 
  select(seqId,perc.identity) %>%
  unique() %>% 
  select(perc.identity) %>% 
  summary()



Fungi.distr.tax.env1 %>% 
  group_by(seqId) %>% 
  mutate(totreads = sum(readcounts)) %>% 
  ungroup() %>% 
  select(seqId, perc.identity, totreads) %>% 
  unique() %>% 
  ggplot + 
  aes(x = perc.identity, 
      fill = totreads) + 
  geom_histogram(bins = 100) + 
  theme_bw() + 
  # xlim(90,100) +
  annotation_custom(gridExtra::tableGrob(Fungi_percid), xmin=80, xmax=90, ymin=6000, ymax=12000) +
  ggtitle("Percentage of identity to the PR2 reference", 
          subtitle= paste0("Fungi, TARA V4; ", 
                           n_distinct(Fungi.distr.tax.env1$seqId), 
                           " distinct amplicons"))

```
</div><div class="column-right">
```{r warning=FALSE, message=FALSE, cache=TRUE}

Fungi.occurrences <- Fungi.distr.tax.env1 %>% 
  group_by(seqId) %>% 
  summarise(Occurrence = n_distinct(Station.label)) %>% 
  arrange(Occurrence) 


fungi.table_occurrences <- Fungi.occurrences %>% 
  select(Occurrence) %>% 
  summary()


Fungi.occurrences %>% 
  ggplot() +
  aes(x=Occurrence) +
  geom_histogram(bins = 100) +
  theme_bw() +
  annotation_custom(gridExtra::tableGrob(fungi.table_occurrences), xmin=25, xmax=55, ymin=7500, ymax=10000) +
  ggtitle("Occurrence of sequences along stations", 
          subtitle= paste0("Fungi, TARA V4; ", 
                           n_distinct(Fungi.occurrences$seqId), 
                           " distinct amplicons"))

```
</div>


<br>

<br> 

<br> 


The dataset is then split into two groups according on the percentage of identity (pid) to the PR2 database:

a. group of sequences with pid $\geqslant$ 80

b. group of sequences with pid < 80


```{r warning=FALSE, message=FALSE, cache=TRUE, fig.align='center'}

Fungi.distr.tax.env1.pidgr <- Fungi.distr.tax.env1 %>% 
  mutate(pid.group = case_when(perc.identity >= 80 ~ "high.pid",
                               perc.identity < 80 ~ "low.pid"))


Fungi.split <- split(Fungi.distr.tax.env1.pidgr, Fungi.distr.tax.env1.pidgr$pid.group)
list2env(Fungi.split, .GlobalEnv)

Fungi.distr.tax.env1.pidgr.annot <- Fungi.distr.tax.env1.pidgr %>%
  group_by(pid.group) %>% 
  summarise(amount = n_distinct(seqId)) %>% 
  ungroup()
  
pie(Fungi.distr.tax.env1.pidgr.annot$amount,
labels = c(paste0("Pid \u2265 80: ", n_distinct(high.pid$seqId), " amplicons"),
           paste0("Pid < 80: ", n_distinct(low.pid$seqId), " amplicons")),
  col=c("forestgreen", "deeppink3"),
  main = "Fungi: Percentage of identity to PR2")



n_distinct(high.pid$seqId) + n_distinct(low.pid$seqId) == n_distinct(Fungi.distr.tax.env1$seqId)

```


The sequences in fasta format are then extracterd for the high percentage group and blasted against a cleaned reference database of dinoflagellates.


```{r warning=FALSE, message=FALSE, cache=TRUE, fig.align='center'}

fungi.high.pid.seqid <- high.pid %>% 
  select(seqId) %>% 
  unique()

write.csv(fungi.high.pid.seqid, "fungi.high.pid.seqid.csv", row.names = F)

###levo il nome della colonna su excel e esporto il file come txt file e lo porto su falkor 
 
###rsync -avz --progress /drives/c/Users/Userszn/Documents/PhD/Git/bin/Pseudonitzschia/fungi.high.pid.seqid.txt lcampese@193.205.231.59:/home/lcampese/Data/metaB/V4/Parasites/Fungi/
```

```{r, eval=FALSE}

###FALKOR
###take fastas

#wc -l fungi.high.pid.seqid.txt: 19406 

##convert file (ascii with crlf terminators) into a simple ascii file, in order to be grepped with the fasta file (simple ascii):

cat fungi.high.pid.seqid.txt | tr -d '\r' > fungi.high.pid.seqid.ascii.txt

#use seqIds  to retrieve fasta of the sequences
srun grep -w -A 1 -Ff fungi.high.pid.seqid.ascii.txt /home/lcampese/Data/metaB/V4/18S_V4_1191_samples.fas --no-group-separator > 18S_V4_1191_samples.fungi.high.pid.fas &
  
#check if it retrieved the right number of sequences

grep ">" 18S_V4_1191_samples.fungi.high.pid.fas -c
#19406
wc -l fungi.high.pid.seqid.ascii.txt
#19406

#ok

##I now have both the fasta: 18S_V4_1191_samples.fungi.high.pid.fas
##and the table with taxonomy, env.par etc: high.pid
```


```{r, eval=FALSE}
#Falkor:

#1)	Trasformo il dataset Dino in formato database 
makeblastdb -in dino.fasta -dbtype nucl -parse_seqids -out dino_db

#3)	Blast con max_target_seqs 1 (mi serve solo il best hit)
blastn -db dino_db -perc_identity 97 -query 18S_V4_1191_samples.fungi.high.pid.fas -outfmt 6 -max_target_seqs 1 -num_threads 8 -out fungi.highpid.checkdino.txt &

#output: wc -l: fungi.highpid.checkdino.txt 
#1223
```



```{r warning=FALSE, message=FALSE, cache=TRUE, fig.align='center'}

fungi.highpid.checkdino <- fread("C:/Users/Userszn/Documents/PhD/script/Pseudonitzschia/metaB/V4/Parasites/Fungi/fungi.highpid.checkdino.txt", header = F)

fungi.highpid.checkdino_seqIds <- fungi.highpid.checkdino %>% 
  select(V1) %>% 
  separate(V1, into = c("seqId", "size"), sep = ";") %>% 
  select(seqId)


high.pid.annot <- high.pid %>% 
  mutate(tax = ifelse(seqId %in% fungi.highpid.checkdino_seqIds$seqId, "dino", "nondino")) %>% 
  select(seqId, perc.identity, tax) %>% 
  group_by(tax) %>% 
  summarise(amount = n_distinct(seqId)) %>% 
  ungroup()


pie(high.pid.annot$amount,
labels = c(paste0("Assigned to dinoflagellates:
", n_distinct(fungi.highpid.checkdino_seqIds$seqId), " amplicons"),
                                   paste0("Not assigned to dinoflagellates:
", (n_distinct(high.pid$seqId) - n_distinct(fungi.highpid.checkdino_seqIds$seqId)), " amplicons")),
    col=c("yellowgreen", "darkgreen"),
    main = "Actual taxonomy of Fungi with pid to PR2 \u2265 80 ")
```

The blast results show that 1223 out of the initial 19406 sequences taxonomically assigned to Fungi are actually belonging to dinoflagellates.

<br>

The dataset is being cleaned by removing the misassignations and sequences in fasta format are then retrieved.

```{r warning=FALSE, message=FALSE, cache=TRUE, fig.align='center'}
high.pid.cleaned <- high.pid %>% 
  filter(! seqId %in% fungi.highpid.checkdino_seqIds$seqId)


n_distinct(high.pid$seqId) - n_distinct(fungi.highpid.checkdino_seqIds$seqId) == n_distinct(high.pid.cleaned$seqId)


high.pid.cleaned.seqid <- high.pid.cleaned %>% 
  select(seqId) %>% 
  unique()

write.csv(high.pid.cleaned.seqid, "high.pid.cleaned.seqid.csv", row.names = F)

###levo il nome della colonna su excel e esporto il file come txt file e lo porto su falkor 
 
###rsync -avz --progress /drives/c/Users/Userszn/Documents/PhD/Git/bin/Pseudonitzschia/high.pid.cleaned.seqid.txt lcampese@193.205.231.59:/home/lcampese/Data/metaB/V4/Parasites/Fungi/
```

```{r, eval=FALSE}
###FALKOR
###take fastas

#wc -l high.pid.cleaned.seqid.txt:  18183

##convert file (ascii with crlf terminators) into a simple ascii file, in order to be grepped with the fasta file (simple ascii):

cat high.pid.cleaned.seqid.txt | tr -d '\r' > high.pid.cleaned.seqid.ascii.txt

#use seqIds  to retrieve fasta of the sequences
srun grep -w -A 1 -Ff high.pid.cleaned.seqid.ascii.txt /home/lcampese/Data/metaB/V4/18S_V4_1191_samples.fas --no-group-separator > 18S_V4_1191_samples.fungi.highpid.cleaned.fas &
  
#check if it retrieved the right number of sequences

grep ">" 18S_V4_1191_samples.fungi.highpid.cleaned.fas -c
#18183
wc -l high.pid.cleaned.seqid.ascii.txt
#18183

#ok

#I now have both the fasta: 18S_V4_1191_samples.fungi.highpid.cleaned.fas
##and the table with taxonomy, env.par etc: high.pid.cleaned
```

The final dataset is made of **18183 sequences** .



<br>

<br> 

Did they include BioMarks sequences in PR2 dataset? 


1. Download supplementary material from the paper [Richards et al.,2015](http://dx.doi.org/10.1098/rspb.2015.2243  "paper BioMarKs").


2. Check headers:

```{r eval=FALSE}
#Falkor: /home/lcampese/Data/metaB/V4/Parasites/Fungi/BioMarKs
#1) combine the 71 fasta files containing sequences of the 71 OTU clusters of the paper into one single fasta files

cat Cluster* > BioMaRks.fas

#take headers of :
# 18S_V4_1191_samples.fungi.highpid.cleaned.fas
# BioMaRks.fas

cat 18S_V4_1191_samples.fungi.highpid.cleaned.fas | grep "^>" | sort > 18S_V4_1191_samples.fungi.highpid.cleaned.fas.header.txt

cat BioMaRks.fas | grep "^>" | sort > BioMaRks.fas.header.txt

##rsync to local
```


```{r warning=FALSE, message=FALSE, cache=TRUE, fig.align='center'}

mine <- fread("C:/Users/Userszn/Documents/PhD/script/Pseudonitzschia/metaB/V4/Parasites/Fungi/biomarks/marine_fungi-v1.0/71_OTU_sequence_clusters/18S_V4_1191_samples.fungi.highpid.cleaned.fas.header.txt", 
              header = F)

biomarks <- fread("C:/Users/Userszn/Documents/PhD/script/Pseudonitzschia/metaB/V4/Parasites/Fungi/biomarks/marine_fungi-v1.0/71_OTU_sequence_clusters/BioMaRks.fas.header.txt",
                  header = F)



mine2 <- mine %>% 
  select(V1) %>% 
  separate(V1, into = c("bla", "seqId"), sep = ">") %>% 
  select(seqId) %>% 
  arrange()


biomarks2 <- biomarks %>% 
  separate(V1, into = c("bla", "ble"), sep = ">") %>% 
  separate(ble, into = c("seqId", "bli"), sep = "_") %>% 
  select(seqId) %>% 
  arrange()


inner_join(mine2, biomarks2)
```

There is no overlap between headers

3. Blast?

```{r, eval=FALSE}
#Falkor: /home/lcampese/Data/metaB/V4/Parasites/Fungi/BioMarKs

#1)	Trasformo il mio dataset PR2 (contenente sequenze di Fungi con perc.identity di almeno 80 e senza le sequenze di Dino) in formato database 
makeblastdb -in ~/Data/metaB/V4/Parasites/Fungi/18S_V4_1191_samples.fungi.highpid.cleaned.fas -dbtype nucl -parse_seqids -out 18S_V4_1191_samples.fungi.highpid.cleaned_db

#3)	Blast con max_target_seqs 1 (mi serve solo il best hit)
blastn -db 18S_V4_1191_samples.fungi.highpid.cleaned_db -perc_identity 100 -query BioMaRks.fas -outfmt 6 -max_target_seqs 1 -num_threads 8 -out fungi.highpid.cleaned.checkBioMarKs.txt &

#wc -l fungi.highpid.cleaned.checkBioMarKs.txt 45 
  
#che vor dì?
```









<br>

To be done:

<br>

Let's look at the low.pid group and check manually if some sequence with a percentage of identity around 75% are actually Fungi