---
title: "Abundance and Distribution of Pseudo-nitzschia in the Global Epipelagic Ocean"
author: "Lucia Campese"
date: "Feb 2020"
output:
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<style>
.column-left{
  float: left;
  width: 50%;
  text-align: left;
}
.column-right{
  float: right;
  width: 50%;
  text-align: right;
}
</style>


## Section One: Pseudo-nitzschia and Total Diatoms' Abundance and Distribution.

### Aim:
The aim is to explore the Pseudo-nizschia OTUs distribution and abundance in relation to the total amount of diatoms along TARA Oceans sampling stations.

<br>

### Dataset
##### Omic data: TARA metaB (V9)
##### Taxon: diatoms (Bacillariophyta)
##### Depth: Surface, Deep Chlorophyll Maximum
##### Size fractions: 3-20um, 5-20um, 20-180um, 180-2000um
##### Geographic scale: Global

<br>

#### Workflow

1. Upload metabarcoding and environmental data, and select depths and size fractions of interest:
+ Depths: surface (SRF) and Deep Chlorophyll Maximum (DCM);
+ Size Fraction: Consider three class:
  a. Nanoplankton: merging 3-20um (Only Arctic) and 5-20um (elsewhere);
  b. Microplankton: 20-180um
  c. Mesoplankton: 180-2000um

```{r, eval=FALSE}

library(tidyverse)
library(ggplot2)

#upload diatom's metab
diatoms <- read.delim(file="C:/Users/Userszn/Google Drive/PhD/TARA_Data/metab/diatoms/diatoms.txt",
                      sep = " ", stringsAsFactors = F) 


#extract Pseudon-itzschia
PN <- diatoms[grep("Pseudo-nitzschia", diatoms$lineage),]


#upload environmental data and filter
dat_samp <- read.delim("~/PhD/TARA_Data/dat_samp.txt") %>% 
  select(Sample.id, Depth, Fraction.size, Station.label,
         Latitude, Longitude, Temperature, Marine.biome, Ocean.region) %>% 
  filter(Depth %in% c("SRF", "DCM"), Fraction.size %in% c("3-20", "5-20", "20-180",
                                                          "180-2000"))


###filter diatoms
diatoms1 <- diatoms[,names(diatoms)%in%dat_samp$Sample.id] %>%
  rownames_to_column("cid") %>%
  gather (Sample.id, values, -cid) %>%
  merge(dat_samp) %>%
  mutate(Fraction.size = case_when(Fraction.size %in% c("3-20", "5-20") ~ "nano",
                                   Fraction.size == "20-180" ~ "micro",
                                   Fraction.size == "180-2000" ~ "meso",
                                   TRUE ~ as.character(Fraction.size)))

#filter pseudonitz
PN1 <- PN[,names(PN)%in%dat_samp$Sample.id] %>% 
  rownames_to_column("cid") %>% 
  gather (Sample.id, values, -cid) %>%
  merge(dat_samp) %>% 
  mutate(Fraction.size = case_when(Fraction.size %in% c("3-20", "5-20") ~ "nano",
                                   Fraction.size == "20-180" ~ "micro",
                                   Fraction.size == "180-2000" ~ "meso",
                                   TRUE ~ as.character(Fraction.size)))

```

<br>

2. Look at the geographic distribution of Pseudo-nizschia OTUs and their contribute to the total diatom abundance per each station.

```{r, eval=FALSE}

PN_diatoms <- diatoms1 %>% 
  filter(values > 0) %>% 
  rename(dia.values = values) %>% 
  mutate(pseu.values = ifelse(cid %in% PN1$cid, dia.values, 0)) %>%
  group_by(Station.label, Fraction.size) %>% 
  mutate(rel.abundance = sum(pseu.values)/sum(dia.values)*100,
         magnitude_overall = factor(10^(ceiling(log10(sum(pseu.values))))),
         magnitude = case_when(magnitude_overall == 100 ~ "1e+02",
                               magnitude_overall == 1000 ~ "1e+03",
                               magnitude_overall == 10000 ~ "1e+04",
                               TRUE ~ as.character(magnitude_overall))) %>%
  ungroup() %>% 
  select(Station.label, magnitude, rel.abundance, Latitude, Longitude, Fraction.size, Depth) %>% 
  filter(rel.abundance > 0) %>% 
  filter(magnitude > 10) %>% 
  unique()

library(maps)
world_map <- map_data("world")
p <- ggplot() + coord_fixed()

base_world3 <- p + geom_polygon(data=world_map,aes(x=long, y=lat,group=group),colour="gray80",size=0.2, fill="gray80")+
  theme_bw() + coord_fixed()

pdf("pseudo-nitzschia_ab_on_diatoms.pdf", width = 8, height = 8)
base_world3+
  geom_point(data=PN_diatoms,
             aes(x = Longitude, y = Latitude, 
                 size = rel.abundance, 
                 fill= magnitude),
             shape=21, col="gray44") +
  theme_minimal() +
  scale_fill_viridis_d() +
  theme(legend.text=element_text(size=10)) +
  scale_size(name="Relative abundance (%)") +
  ggtitle("Pseudo-nitzschia abundance expressed as 
          percentage on total diatom reads per station")
dev.off()

```

![pseudo-nitzschia_ab_on_diatoms](C:/Users/Userszn/Google Drive/PhD/script/Pseudonitzschia/metaB/pseudo-nitzschia_ab_on_diatoms.pdf){width=100% height=700}

<br>

3. Look at the correlation between Pseudo-nitzschia OTUs and total diatom abundance.

```{r eval=FALSE}
##Pearson correlation
Pearson_cor <- diatoms1 %>% 
  filter(values > 0) %>% 
  rename(dia.values = values) %>% 
  mutate(pseu.values = ifelse(cid %in% PN1$cid, dia.values, 0)) %>%
  group_by(Station.label) %>% 
  summarise(dia.sum = sum(dia.values),
            pseu.sum = sum(pseu.values)) %>%
  ungroup() %>% 
  arrange(Station.label) %>% 
  filter(dia.sum > 0, pseu.sum > 0)

cor.test(log(Pearson_cor$dia.sum), log(Pearson_cor$pseu.sum), method = "pearson")

```

```{r, eval=FALSE}
pdf("pn_dia_cor.pdf")
diatoms1 %>% 
  filter(values > 0) %>% 
  rename(dia.values = values) %>% 
  mutate(pseu.values = ifelse(cid %in% PN1$cid, dia.values, 0)) %>%
  group_by(Station.label, Fraction.size) %>% 
  summarise(dia.sum = sum(dia.values), pseu.sum = sum(pseu.values)) %>%
  ungroup() %>% 
  ggplot() + 
  geom_point(aes(x = dia.sum, y = pseu.sum)) +
  theme_minimal() +
  scale_x_log10() +
  scale_y_log10() +
  theme(legend.text=element_text(size=10)) +
  xlab("total diatom reads") + 
  ylab("total Pseudo-nitzschia reads") +
  ggtitle("Pseudo-nitzschia and total diatom reads per station") +
  annotate("text", x=10, y=100000, label = 
"Pearson's correlation
cor 0.797
p-value < 2.2e-16
95% ci. [0.728; 0.850]")
dev.off()
```

![pseudo-pn_dia_cor](C:/Users/Userszn/Google Drive/PhD/script/Pseudonitzschia//metaB/pn_dia_cor.pdf){width=100% height=700}

<br>

```{r eval=FALSE}
##lm
prova <- diatoms1 %>% 
  filter(values > 0) %>% 
  rename(dia.values = values) %>% 
  mutate(pseu.values = ifelse(cid %in% PN1$cid, dia.values, 0)) %>%
  group_by(Station.label) %>% 
  summarise(dia.sum = sum(dia.values), pseu.sum = sum(pseu.values),
            y = pseu.sum/dia.sum) %>%
  ungroup() %>% 
  rename(x=dia.sum, y0=pseu.sum) %>%
  filter(y0>0) %>% 
  mutate(x=log(x), y=log(y)) %>% 
  rownames_to_column("id") %>% 
  select(-c("Station.label"))

model1 <- lm(y~x, data=prova)
temp_var <- predict(model1, interval="prediction")
new_df <- cbind(prova,temp_var)
p <- ggplot(new_df) + aes(x,y) +
  geom_label(aes(label=id)) +
  geom_line(aes(y=lwr), color = "red", linetype = "dashed")+
  geom_line(aes(y=upr), color = "red", linetype = "dashed")+
  geom_smooth(method=lm, se=TRUE) +
  theme_minimal()
print(p)
par(mfrow=c(2,3)); plot(model1, which=1:6)

```

<br>
<br>

## Section Two: Within-genus exploration

### Aim:
The aim is to look more in detail at Pseudo-nitzschia OTUs distribution and to explore the number and proportion of OTUs taxonomically assigned at species level

<br>

### Dataset
##### Omic data: TARA metaB (V9)
##### Taxon: Pseudo-nitzschia
##### Depth: Surface, Deep Chlorophyll Maximum
##### Size fractions: all
##### Geographic scale: Global  

<br>

#### Workflow
1. Check for spatial autocorrelation.

```{r, eval=FALSE}
library(tidyverse)

##autocorrelation
acf <- PN1 %>%
  mutate(Station.label = substr(Station.label, 6, 8)) %>% 
  filter(values > 0) %>% 
  group_by(Station.label) %>% 
  mutate(sum = sum(values)) %>% 
  ungroup() %>%
  select(Station.label, sum) %>%
  arrange(Station.label) %>% 
  unique()

pdf("autocorrelation.pdf")
acf(acf$sum)
dev.off()
```

![pseudo-autocorrelation](C:/Users/Userszn/Google Drive/PhD/script/Pseudonitzschia//metaB/autocorrelation.pdf){width=100% height=700}

<br>

2. Look at the distribution of Pseudo-nitzschia OTUs both in terms of single values and sum of values per each station.


<div class="column-left">
```{r, eval=FALSE}
##histogram
##scatterplot temperature & n reads
pdf("pn_ab_n_reads.pdf")
PN1 %>%
  mutate(Station.label = substr(Station.label, 6, 8)) %>% 
  filter(values !=0) %>% 
  ggplot() +
  geom_histogram(aes(values), alpha=0.5) +
  theme_minimal() +
  scale_x_log10() +
  xlab("abundance") +
  ggtitle("Pseudo-nitzschia abundance expressed as number of reads")
dev.off()
```
![pseudo-pn_ab_n_reads](C:/Users/Userszn/Google Drive/PhD/script/Pseudonitzschia/metaB/pn_ab_n_reads.pdf){width=100% height=700}</div>

<div class="column-right">
```{r, eval=FALSE}
pdf("pn_ab_n_reads_station.pdf")
PN1 %>%
  mutate(Station.label = substr(Station.label, 6, 8)) %>% 
  group_by(Station.label) %>% 
  mutate(sum = sum(values)) %>% 
  ungroup() %>%
  ggplot() +
  geom_histogram(aes(sum), alpha=0.5) +
  theme_minimal() +
  scale_x_log10() +
  xlab("abundance per station") +
    ggtitle("Pseudo-nitzschia abundance expressed as number of reads per station")
dev.off()
```
![pseudo-pn_ab_n_reads_station](C:/Users/Userszn/Google Drive/PhD/script/Pseudonitzschia/metaB/pn_ab_n_reads_station.pdf){width=100% height=700}</div>

<br>

3. Look at the geographic distribution of Pseudo-nizschia OTUs.

```{r, eval=FALSE}

global_maps <- PN1 %>%
  mutate(Station.label = substr(Station.label, 6, 8)) %>% 
  group_by(Station.label) %>% 
  mutate(sum = sum(values)) %>% 
  filter(sum > 10) %>% 
  mutate(magnitude_overall = factor(10^(ceiling(log10(sum)))),
         magnitude = case_when(magnitude_overall  == "100" ~ "1e+02",
                               magnitude_overall == "1000" ~ "1e+03",
                               magnitude_overall == "10000"  ~ "1e+04",
                               TRUE ~ as.character(magnitude_overall)))

pdf("pn_ab_n_reads_station_map.pdf")
base_world3+
  geom_point(data= global_maps,
             aes(x = Longitude, y = Latitude,
                 size = magnitude, fill = magnitude),
             shape=21, col="gray44") +
  theme_minimal() +
  theme(legend.text=element_text(size=10)) +
  scale_fill_brewer(palette = "Paired") +
  ggtitle("Pseudo-nitzschia abundance expressed as number of reads per station")
dev.off()
```

![pn_ab_n_reads_station_map](C:/Users/Userszn/Google Drive/PhD/script/Pseudonitzschia/metaB/pn_ab_n_reads_station_map.pdf){width=100% height=700}

<br>

4. Look at the distribution of Pseudo-nitzschia OTUs along the Temperature gradient.

```{r, eval=FALSE}
library(ggforce)

pdf("pn_ab_reads_station_temperature.pdf")
PN1 %>%
  mutate(Station.label = substr(Station.label, 6, 8)) %>% 
  filter(values > 0) %>% 
  group_by(Station.label) %>% 
  mutate(sum = sum(values)) %>% 
  ungroup() %>% 
  select(Station.label, Temperature, sum, values) %>% 
  unique() %>% 
  mutate(group = ifelse(Temperature <= 11.5, "cold", "warm")) %>% 
  drop_na() %>%
  ggplot() +
  geom_point(aes(x = Temperature, y = sum)) +
  scale_y_log10() +
  theme_minimal() +
  ylab("number of reads") +
  theme(legend.text=element_text(size=10)) +
  ggtitle("Pseudo-nitzschia abundance expressed as number of reads per station") +
  geom_mark_ellipse(aes(x = Temperature, y = sum, fill=group, label = group),
                    show.legend = F,
                    con.cap = 0,
                    expand = unit(0.2, "cm")) +
  scale_fill_manual(values = c("#00BFC4", "#F8766D")) +
  annotate("text", x = 20, y = 7.5e+05, label = "
           Min.:  143
           Median: 6519  
           Mean: 76096  
           Max.: 706423 ") +
  annotate("text", x = 2, y = 23, label = "
           Min.:  1
           Median: 270  
           Mean: 3902  
           Max.: 179138 ") 
dev.off()
```

![pn_ab_reads_station_temperature](C:/Users/Userszn/Google Drive/PhD/script/Pseudonitzschia/metaB/pn_ab_reads_station_temperature.pdf){width=100% height=700}

<br>

5. Explore the difference in terms of size fraction

```{r, eval=FALSE}

###Venn
library(venn)
pdf("size_fr_venn.pdf")
PN1 %>%
  group_by(cid, Fraction.size) %>% 
  summarise(sum = sum(values)) %>%
  ungroup() %>% 
  filter(sum > 0) %>% 
  mutate(sum = ifelse(sum > 0,  1, 0)) %>%
  spread(Fraction.size, sum) %>% 
  mutate_at(vars(-cid), funs(ifelse(is.na(.), 0, .))) %>% 
  select(-cid) %>% 
  venn(ellipse = F, borders = F,  cexil = 0.8, cexsn = 0.9)
dev.off()
```
![size_fr_venn](C:/Users/Userszn/Google Drive/PhD/script/Pseudonitzschia/metaB/size_fr_venn.pdf){width=100% height=700}


<br>

6. Look at the geographic distribution of Pseudo-nizschia OTUs relative to Size and Depth.

```{r, eval=FALSE}

pdf("pn_ab_each_size_fr.pdf")
##map of abundance per each size
base_world3+
  geom_point(data=global_maps, aes(x = Longitude, y = Latitude,
                                   size = magnitude, fill = magnitude),
             shape=21, col="gray44") +
  theme_minimal() +
  theme(legend.text=element_text(size=10)) +
  scale_fill_brewer(palette = "Paired") +
  ggtitle("Pseudo-nitzschia abundance expressed as number of reads per station") + 
  facet_wrap(~Depth + Fraction.size) 
dev.off()
```

![pn_ab_each_size_fr](C:/Users/Userszn/Google Drive/PhD/script/Pseudonitzschia/metaB/pn_ab_each_size_fr.pdf){width=100% height=700}

<br>

7. Look at the genus richness in terms of different OTUs and look at the geographic distibution and abundance of each of the 8 species identified.

```{r, eval=FALSE}

load("C:/Users/Userszn/Google Drive/PhD/TARA_Data/metaB/euk_full_names.Rdata")
library(tidyverse)


taxonomy <- euk_full_names %>% 
  group_by(OTU, rank) %>%
  summarise(name =  paste(name, collapse = ";")) %>%
  ungroup %>%
  spread(rank, name) %>% 
  filter(genus == "Pseudo-nitzschia") %>% 
  select(OTU, genus, species) %>% 
  mutate_at(vars(species), funs(ifelse(is.na(.), "unknown"  , .)))


global_maps_taxonomy <-  global_maps %>% 
  rename(OTU=cid) %>% 
  merge(taxonomy)

library(stringr)
library(treemapify)

pdf("pn_taxonomy_treemap.pdf")
global_maps_taxonomy %>% 
  mutate(species = str_replace(species, "Pseudo-nitzschia", "P.")) %>%
  group_by(species) %>% 
  summarise(richness = n_distinct(OTU)) %>% 
  ungroup() %>%
  ggplot(aes(area=richness, fill=species, label=species)) +
  geom_treemap(show.legend = T) + 
  geom_treemap_text(fontface = "italic",
                    color = "white",
                    place = "center", show.legend = F) +
  ggtitle("", 
          subtitle = paste0(n_distinct(global_maps_taxonomy$OTU)," OTUs"))
dev.off()
```
![pn_taxonomy_treemap](C:/Users/Userszn/Google Drive/PhD/script/Pseudonitzschia/metaB/pn_taxonomy_treemap.pdf){width=100% height=700}



```{r, eval=FALSE}

library(scatterpie)

vediamo <- global_maps_taxonomy %>%
  filter(values > 0) %>% 
  group_by(Station.label) %>% 
  mutate(Longitude=mean(Longitude), Latitude = mean(Latitude)) %>% 
  ungroup() %>% 
  group_by(Station.label, species) %>% 
  mutate(fraction = sum(values)/sum) %>% 
  ungroup() %>% 
  select(Station.label, Longitude, Latitude, species, fraction, sum) %>% 
  arrange(Station.label) %>% 
  unique()  %>% 
  mutate(species = str_replace(species, "Pseudo-nitzschia", "P.")) %>% 
  group_by(Station.label) %>% 
  mutate(radius = max(fraction)*4) %>% 
  spread(species, fraction) %>% 
  mutate_all(funs(ifelse(is.na(.), 0, .)))

pdf("pn_species_distribution_map.pdf", width = 10)
base_world3 + 
  geom_scatterpie(aes(x=Longitude, y=Latitude, r=radius),
                  data=vediamo,
                  cols= c("P. australis", "P. calliantha", "P. delicatissima",
                          "P. fraudulenta", "P. heimii", "P. multiseries",
                          "P. pseudodelicatissima", "P. pungens", "unknown"),
                  alpha=.7) +
  theme_minimal() +
  theme(legend.text=element_text(size=10)) +
  coord_fixed() +
  scale_color_discrete() +
  ggtitle("Pseudo-nitzschia abundance and distribution at species level") #+
  #geom_scatterpie_legend(vediamo$radius, x=-150, y=-70)
dev.off()

##vedi Bates 2018 Fig. 2
```
![pn_species_distribution_map](C:/Users/Userszn/Google Drive/PhD/script/Pseudonitzschia/metaB/pn_species_distribution_map.pdf){width=100% height=700}